---
title: "Olimi"
output: html_document
date: "2025-05-27"
---


microbiome")

#replace this with your working directory
```{r, echo=F, warning= F}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
library(phyloseq);packageVersion("phyloseq")#‘1.46.0’
library(vegan);packageVersion("vegan")# ‘2.6.4’
library(genefilter);packageVersion("genefilter")#‘1.84.0’
library(psych);packageVersion("psych")#‘2.4.1’
library(DESeq2);packageVersion("DESeq2")#‘1.42.1’
library(reshape2);packageVersion("reshape2")#‘1.4.4’
library(microbiome);packageVersion("microbiome")#‘1.24.0’
library(ranacapa);packageVersion("ranacapa")#‘0.1.0’
library(S4Vectors);packageVersion("S4Vectors")#‘0.40.2’
library(Biobase);packageVersion("Biobase")#‘2.62.0’
library(BiocGenerics);packageVersion("BiocGenerics")#‘0.48.1’
library(parallel);packageVersion("parallel")#‘4.3.2’
library(ggpubr);packageVersion("ggpubr")#‘0.6.0’
#library(metagMisc)
library(decontam);packageVersion("decontam")#‘1.22.0’
library(RColorBrewer);packageVersion("RColorBrewer")#‘1.1.3’
library(multcompView);packageVersion("multcompView")#‘0.1.10’
library(metagenomeSeq);packageVersion("metagenomeSeq")#‘1.43.0’
#library(devtools)
library(lawstat);packageVersion("lawstat")#‘3.6’
library(DirichletMultinomial);packageVersion("DirichletMultinomial")#‘1.44.0’
library(lattice);packageVersion("lattice")# ‘0.22.5’
library(xtable);packageVersion("xtable")#‘1.8.4’
library(permute);packageVersion("permute")#‘0.9.7’
library(grid);packageVersion("grid")#‘4.3.2’
library("doParallel");packageVersion("doParallel")#‘1.0.17’
#library(igraph)
library(ade4);packageVersion("ade4")#‘1.7.22’
#library(rgl)
library(car);packageVersion("car")#‘3.1.2’
library(gridExtra);packageVersion("gridExtra")# ‘2.3’
library(tidyverse);packageVersion("tidyverse")#‘2.0.0’
library(phia);packageVersion("phia")#‘0.3.1’
library(lme4);packageVersion("lme4")#‘1.1.35.1’
library(broom);packageVersion("broom")#‘1.0.5’
library(AICcmodavg);packageVersion("AICcmodavg")#‘2.3.3’
library(multcomp);packageVersion("multcomp")#‘1.4.25’
library(rcompanion);packageVersion("rcompanion")#‘2.4.35’
library(hrbrthemes);packageVersion("hrbrthemes")#‘0.8.7’
library(viridis);packageVersion("viridis")#‘0.6.5’
library(ape);packageVersion("ape")#‘5.7.1’#loading the tree
library(microbiomeutilities);packageVersion("microbiomeutilities")#‘1.0.17’#process taxonomy table
library(microViz);packageVersion("microViz")#‘0.12.1’#clean taxonomy table to be cited
library(GUniFrac);packageVersion("GUniFrac")#‘1.8’
library(ggtext) ;packageVersion("ggtext")# ‘0.1.2’for rotated labels on ord_plot() 
library(ggraph);packageVersion("ggraph")# ‘2.2.1’for taxatree_plots()
library(corncob);packageVersion("corncob")#‘0.4.1’ for beta binomial models in tax_model()
library(VennDiagram);packageVersion("VennDiagram")#‘1.7.3’
library(UpSetR);packageVersion("UpSetR")#‘1.4.0’
library(rstatix);packageVersion("rstatix")#‘0.7.2’
library(ALDEx2);packageVersion("ALDEx2")#[1] ‘1.34.0’
library(cowplot);packageVersion("cowplot")#‘1.1.3’
library(HMP);packageVersion("HMP")#HMP: Hypothesis Testing & Power Calculations for Comparing Metagenomic Samples from HMP #‘2.0.1’
library(dendextend);packageVersion("dendextend")#‘1.17.1’
library(microViz)
library(ggtext)
library(ggraph)
library(DT)
library(corncob)
library(patchwork)
library(ggplot2) # 3.2.1
library(dplyr)
library(lubridate)
library(cowplot) # 1.0.0
library(egg) # 0.4.5
library(microbiome)
library(knitr)
library(tidyplots)
library(rcompanion)
library(multcompView)
library(tidyverse)
library(ggpubr)
library(dunn.test)
library(FSA)
#install.packages("explore")
library(explore)
library(gcookbook)
library(tidyverse)
library(patchwork)
library(MicrobiotaProcess)
library(rmarkdown)#install.packages("rmarkdown")
library(dada2)
library(Biostrings)
library(ggalluvial)
library(ggh4x)
library(tidyverse)
library(ggrepel)
library(qiime2R)#https://github.com/jbisanz/qiime2R
library(pals)
library(ggforce)
library(gridExtra)
library(vegan)
library(data.table)
library(picante)
library(ggtreeExtra)
library(ggpmisc)
library(DECIPHER) #BiocManager::install("DECIPHER")
library(phangorn) #remotes::install_github("KlausVigo/phangorn")
library(ggtree) #BiocManager::install("ggtree")
library(plotly)
library(speedyseq) #remotes::install_github("mikemc/speedyseq")
library(ComplexHeatmap)
library(agricolae)
library(scales)
library(ggpmisc)
library(pairwiseAdonis)#library(devtools) install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
#remotes::install_github("mahendra-mariadassou/phyloseq-extended", ref = "dev")
library(phyloseq.extended)#remotes::install_github(repo = "mahendra-mariadassou/phyloseq-extended@v0.1.0.9000-beta", ref = "dev")
#The following need to be installed but NOT loaded (same naming of different functions):
#library(metagMisc) #devtools::install_github("vmikk/metagMisc")
#library(metagenomeSeq) #BiocManager::install("metagenomeSeq")
source("Rscript_parwise.adonis.txt")
source("aggregate_top_taxa.txt")
source("PD.txt")
source("aggregate_rare.txt")
source("plot_composition.txt")
source("RSCRIPT_parwise.adonis.txt")
source("plot_hierarchy.R")
#https://sites.northwestern.edu/researchcomputing/resources/using-git-&-github-with-r-rstudio/#github-r config
```



#Beta diversity
#CSS transformation
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
phyloseq_transform_css <- function(physeq, norm = TRUE, log = TRUE, ...){
  # require(metagenomeSeq)
  MGS <- phyloseq::phyloseq_to_metagenomeSeq(physeq)
  # get the normalized count matrix
  otu_norm <- metagenomeSeq::MRcounts(MGS, norm = norm, log = log, ...)
  # exportMat(datt.norm, file = "tmp.txt")    # norm = TRUE, log = TRUE
  ## save sample statistics (sample scaling factor, quantile value, number of identified features and library size):
  # exportStats(dattM.sf, file = "tmp_stats.txt")
  # Substitue raw abundance to the css-normalized data
  physeq.tr <- physeq
  phyloseq::otu_table(physeq.tr) <- phyloseq::otu_table(otu_norm, taxa_are_rows = T)
  return(physeq.tr)
}
```



#Load colors for identifying the different samples
```{r,echo = F,warning = FALSE}
mycolors=c("grey35","#00AFBB","goldenrod","#0dba2f","#392682","coral3","#8E9CA3",
           "chartreuse4","grey85","#FF2700","bisque4","#0B775E","darkred","#F39B7FB2","#FA41CA",
           "steelblue","blue","#D16103", "#C3D7A4", "#52854C","deepskyblue","#293352","#FFDB6D",
           "Yellow","#CC79A7","#392682", "#CAB2D6","maroon","orchid1","khaki2","#86486F","#7E6148B2");mycolors

mycola=c("grey35","chartreuse4","#FE7F9D","darkturquoise", "#F0E442","#009E73","#FF0000",
         "Orange","blue","#850dba","#FF6AD5","#6A3D9A","Yellow","Blue",
         "deeppink","#1C0221","#00AFBB","steelblue","#D16103", "#C3D7A4","#52854C",
         "#4E84C4","#293352","#FFDB6D","#CC79A7","#C4961A", "#F4EDCA","dodgerblue2", 
         "#E31A1C","green4","#0dba2f","black","gold1", "skyblue2","#FB9A99","#CAB2D6",
         "blue1","steelblue4","Sky Blue","grey85","#3B9AB2","goldenrod","yellow3",
         "darkorange4","brown","#999999", "#E69F00","#56B4E9","#009E73","#F0E442",
         "#D55E00", "#E69F00","#ba0d42","yellow4","#D55E00", "#CC79A7","darkred",
         "#56B4E9","#86486F","#446455", "bisque4","#F7B1AB","#807182","#E3DDDD",
         "#A45C3D","coral3","#000099","#E2C59F","#B6C5CC", "#8E9CA3","#556670",
         "#771C19", "#AA3929","#E25033","#F27314", "#F8A31B","#2D2D2D","#91D1C2B2",
         "#DC0000B2","#7E6148B2","#E64B35B2", "#4DBBD5B2","#00A087B2","#3C5488B2",
         "#F39B7FB2","#8491B4B2","#FF2700","#62BBA5","#B58900","#80CDC1","#8073AC",
     "#0072B2","#392682","#FA41CA","#0B775E","#02401B","#F4B5BD","#B6854D","deepskyblue","green1");mycola

mycol = c("bisque4", "deeppink", "chartreuse4","blue","coral3","goldenrod","deepskyblue","darkred","#0dba2f", "Orange","Yellow", "#850dba","Blue", "Sky Blue", "#ba0d42");mycol
```

#Bacterial community



#Load data (ASV table, taxonomy & tree, & metadata_bac)
```{r,echo = F}
##Imported feature table
asv_tab_bac <- read.table("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Bacteria_Table.txt", 
              header=TRUE,
              row.names="OTUID");asv_tab_bac
asv_tab_bac <- otu_table(asv_tab_bac, taxa_are_rows = TRUE);asv_tab_bac 

## Imported taxonomy table
tax_tab_bac <- read.table("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Bacteria_Taxonomy.txt", 
              header=TRUE,
              sep="\t",
              row.names="OTUID");tax_tab_bac
tax_tab_bac <- as.matrix(tax_tab_bac);tax_tab_bac
tax_tab_bac <- tax_table(tax_tab_bac);tax_tab_bac

## Imported metadata_bac file
metadata_bac <- read.csv("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Bacteria_Metadata.csv",
              header=TRUE,
              row.names="SampleID");metadata_bac
#rownames(metadata_bac) <- metadata_bac[,1]
metadata_bac <- sample_data(metadata_bac);metadata_bac

#import reference sequences
## Import reference sequences
# rs_table <- Biostrings::readDNAStringSet(
#   "outputs/qiime2/dna-sequences.fasta",
#   format = "fasta",
#   nrec = -1L,
#   skip = 0L,
#   seek.first.rec = F,
#   use.names = T,
#   with.qualities = F)

#proof that the data are of equal dimensions
all(rownames(metadata_bac) %in% colnames(asv_tab_bac))

#Convert to phyloseq object
ps_Bacteria <- merge_phyloseq(
  asv_tab_bac,
  tax_tab_bac, 
  metadata_bac);ps_Bacteria#[ 16460 taxa & 1197 samples ]
summarize_phyloseq(ps_Bacteria)

ps0_Bacteria <- ps_Bacteria # 16460 taxa & 1197 samples
save(ps0_Bacteria,
file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps0_Bacteria.RData")

```


#clean the data
```{r,echo = FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps0_Bacteria.RData")

ps_clean_Bacteria = subset_taxa(ps0_Bacteria, Kingdom=="Bacteria");ps_clean_Bacteria# [ 6448 taxa & 84 samples ]
ps_clean_Bacteria = subset_taxa(ps_clean_Bacteria, Phylum!="Cyanobacteria");ps_clean_Bacteria#  [ 15588 taxa & 1197 samples ]
ps_clean_Bacteria = subset_taxa(ps_clean_Bacteria, Order!="Chloroplast");ps_clean_Bacteria#  [ 15588 taxa & 1197 samples ]
ps_clean_Bacteria = subset_taxa(ps_clean_Bacteria, Family!="Mitochondria");ps_clean_Bacteria# [ 15588 taxa & 1197 samples ]
ps_clean_Bacteria <- prune_taxa(taxa_sums(ps_clean_Bacteria)>0, ps_clean_Bacteria);ps_clean_Bacteria# [ 15588 taxa & 1197 samples ]

taxonomy <- as.vector(phyloseq::tax_table(ps_clean_Bacteria))
sums_ps_clean_Bacteria <- sample_sums(ps_clean_Bacteria);sums_ps_clean_Bacteria
sums_ps_clean_Bacteria <- data.frame(sums_ps_clean_Bacteria);sums_ps_clean_Bacteria
sums_ps_clean_Bacteria <- as.matrix(sums_ps_clean_Bacteria);sums_ps_clean_Bacteria
colSums(sums_ps_clean_Bacteria) #  2684660      
max(sums_ps_clean_Bacteria) #199701
min(sums_ps_clean_Bacteria) # 1889
sums_ps_clean_Bacteria

ps1_Bacteria <- ps_clean_Bacteria

sort(sample_sums(ps1_Bacteria))
summary(sample_sums(ps1_Bacteria)) # mean reads/sample = 22791
summarize_phyloseq(ps1_Bacteria) # Average reads = 22791, Total reads = 27281011, Singletons = 38
taxa_names(ps1_Bacteria)[1:5]
sample_data(ps1_Bacteria)$total_reads <- sample_sums(ps1_Bacteria)

save(ps1_Bacteria, file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps1_Bacteria.RData")

K <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Kingdom"]));K#
P <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Phylum"]));P # 27 phyla
C <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Class"]));C # 53 classes
O <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Order"]));O # 137 orders
Fa <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Family"]));Fa # 251 families
G <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Genus"]));G # 503 genera
#S <- as.data.frame(table(tax_table(ps1_Bacteria)[,"Species"]));S # 1264 species

microbiome::summarize_phyloseq(ps1_Bacteria)
#Number of singletons = 68"
#Sparsity = 0.995179763625127"
#Average number of reads = 22794.9197994987"
# Total number of reads = 27281011"
#Max. number of reads = 225892"
#Min. number of reads = 68"
taxonomy <- as.vector(phyloseq::tax_table(ps1_Bacteria))
sums_ps1_Bacteria <- sample_sums(ps1_Bacteria);sums_ps1_Bacteria
sums_ps1_Bacteria <- data.frame(sums_ps1_Bacteria);sums_ps1_Bacteria
sums_ps1_Bacteria <- as.matrix(sums_ps1_Bacteria);sums_ps1_Bacteria
colSums(sums_ps1_Bacteria) #  27281011      
max(sums_ps1_Bacteria) #225892
min(sums_ps1_Bacteria) # 68
sums_ps1_Bacteria
```


#Filter data at 0.25 prevalence %

#Obtained from: https://f1000research.com/articles/5-1492/v1;#One of the reasons to filter by prevalence is to avoid spending #much time analyzing taxa that were only rarely seen. This also turns out to be a useful filter of noise (taxa that are #actually just artifacts of the data collection process)
# Define prevalence of each taxa; # (in how many samples did each taxa appear at least once)

```{r,echo = FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps1_Bacteria.RData")

prev0 = apply(X = otu_table(ps1_Bacteria),
              MARGIN = ifelse(taxa_are_rows(ps1_Bacteria), yes = 1, no = 2),
              FUN = function(x){sum(x > 0)})

prevdf = data.frame(Prevalence = prev0,
                    TotalAbundance = taxa_sums(ps1_Bacteria),
                    tax_table(ps1_Bacteria))# the most prevalent ASVs present therein the composition

#only for reference, in case we want to filter Phyla that appear less than X times
keepPhyla = table(prevdf$Phylum)[(table(prevdf$Phylum) > 0)]

prevdf1 = subset(prevdf, Phylum %in% names(keepPhyla))

# Define prevalence threshold as 0.25% of total samples# 
prevalenceThreshold = 0.0025 * nsamples(ps1_Bacteria)
prevalenceThreshold#0.075

# Execute prevalence filter, using `prune_taxa()` function
ps2_Bacteria_Bacteria = prune_taxa((prev0 > prevalenceThreshold), ps1_Bacteria);ps2_Bacteria_Bacteria

# Filter entries with unidentified Phylum.
ps3_Bacteria = subset_taxa(ps2_Bacteria_Bacteria, Phylum %in% names(keepPhyla));ps3_Bacteria

#Plot prevalence of different phyla
#https://github.com/joey711/phyloseq/issues/1474

ggplot(prevdf1, aes(TotalAbundance, Prevalence, color = Phylum)) +
  geom_hline(yintercept = prevalenceThreshold, alpha = 0.5, linetype = 2) +
  geom_point(size = 2, alpha = 0.65) +
  scale_y_log10() + scale_x_log10() +
  ggtitle(paste0("Prevalence filtering, using a threshold of ", prevalenceThreshold))+
  xlab("Total Abundance") +
  scale_color_manual(values=ocean.balance(38))+
  theme_bw()+
  facet_wrap(~Phylum, nrow = 4)

print(paste0("Initial #ASVs:", ntaxa(ps1_Bacteria), ". #ASVs after prevalence filtering:", ntaxa(ps3_Bacteria), ". #ASVs removed: ", ntaxa(ps1_Bacteria)-ntaxa(ps3_Bacteria)))
```


# Save the bacterial object
```{r}
save(ps3_Bacteria, file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
```




## Ab-Prev prep
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")

meta <- microbiome::meta(ps3_Bacteria) # x_bac
asv <- otu_table(ps3_Bacteria) # x_bac
taxo <- rownames_to_column(as.data.frame(tax_table(ps3_Bacteria)),
                                        "ASV") # x_bac
sample_names(ps3_Bacteria)
# presence-absence data
asv_PA <- 1 * ((asv > 0) == 1)

# occupancy calculation
asv_prev <- rowSums(asv_PA) / ncol(asv_PA)

# relative abundance
asv_rel <- apply(vegan::decostand(asv, 
                                  method = "total", 
                                  MARGIN = 2), 1, mean)

# combine occupancy and RA of each ASV in a table
asv_prev_rel <- rownames_to_column(as.data.frame(cbind(asv_prev, 
                                                       asv_rel)),
                                   "ASV")

asv_prev_rel_taxo <- merge(taxo, 
                           asv_prev_rel, 
                           by = "ASV")

# merge with metadata
asv_tdf <- rownames_to_column(as.data.frame(t(asv)),
                           "replicates")
library(dplyr)
meta <- meta <- meta %>% rename_with(~ "replicates", 1)
colnames(meta)

colnames(asv_tdf)
asv_tdf_meta <- merge(meta,
                      asv_tdf, 
                      by = "replicates")

# keep only the one feature column for collaps3_Bacteriaing table
asv_tdf_meta <- asv_tdf_meta[-c(1, 3:29)]

asv_gen <- asv_tdf_meta %>% 
    group_by(Genotype) %>%
  summarise_each(list(sum))

# make a ASV P/A table
asv_gen_PA <- 1 * ((asv_gen > 0) == 1)

# merge prevalence, RA and taxa with the nr of obs per genotype
asv_gen_obs <- data.frame(colSums(asv_gen_PA)) %>%
  na.omit() %>% 
  rownames_to_column("ASV")
names(asv_gen_obs)[names(asv_gen_obs) == "colSums.asv_gen_PA."] <- 'Obs'

asv_prev_rel_taxo_gen_obs <- merge(asv_prev_rel_taxo, 
                                   asv_gen_obs, 
                                   by = "ASV")
```

## Ab-Prev Plot 
```{r}
counts <- asv_prev_rel_taxo_gen_obs %>% dplyr::count(Kingdom = "Bacteria")

p_core_flex1_all_bacteria <- 
  ggplot(asv_prev_rel_taxo_gen_obs, 
         aes(x = asv_rel, y = asv_prev, size = Obs)) +
  geom_point(
    color = "darkgrey",
    alpha = 0.5) + 
  ggrepel::geom_text_repel(data = subset(asv_prev_rel_taxo_gen_obs, asv_prev > 0.5),
                           max.overlaps3_Bacteria = 20,
                           aes(asv_rel,
                               asv_prev,
                               label = Genus),
                           size = 3,
                           show.legend = F) +
  geom_text(data = counts,
            aes(x = 0.00000001, y = 0.1, 
                label = paste0("Total ASVs\n", n),  
                fontface = "bold"), size = 4) +
  geom_hline(yintercept = 0.9, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.7, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.5, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.25, linetype = "dotted", color = "red") +
  geom_vline(xintercept = 0.00005, linetype = "dotted", color = "red", show.legend = T) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.line = element_blank(),
    legend.text = element_text(size = 12),
    legend.position = c(0.35, 0.95),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "transparent", color = "transparent"),
    legend.margin = margin(0, 0, 0, 0),
    panel.spacing = unit(0.1, "lines")) +
  scale_x_log10(breaks = c(0.1, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001), 
                labels = c(10, 1, 0.1, 0.01, 0.001, 0.0001, 0.00001)) +
  scale_y_continuous(limits = c(-0.01, 1.01),
                     breaks = c(0,0.25, 0.5, 0.7, 0.9),
                     labels = c(0,25, 50, 70, 90)) +
  labs(x = "Relative abundance [%]", y = "Prevalence [%]", size = "nº cultivars"); p_core_flex1_all_bacteria

 ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ab-prev-all_bacteria.tiff",     
  plot = p_core_flex1_all_bacteria,
   scale = 1,
 width = 25,
  height = 20,
  units = c("cm"),
 dpi = 600)
```


#comsider order level
#https://www.geeksforgeeks.org/networkd3-package-in-r/
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
library("MiscMetabar")
library("phyloseq")
library("networkD3")
#summary_plot_pq(ps3_Bacteria)
#sankey_pq(gp, taxa = c(1:5))
#upset_pq(ps3_Bacteria, "Genotype", taxa = "Genus")

toptax = names(sort(taxa_sums(ps3_Bacteria), decreasing=TRUE))[1:100];toptax
ps.toptax_Bacteria = transform_sample_counts(ps3_Bacteria, function(x) {x/sum(x)*100});ps.toptax_Bacteria
ps.toptax_Bacteria <- prune_taxa(toptax, ps.toptax_Bacteria);ps.toptax_Bacteria
upset_bacteria_Order=upset_pq(ps.toptax_Bacteria, "Genotype", taxa = "Order");upset_bacteria_Order

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Shared_upset_bacteria_Order.tiff",     
  plot = upset_bacteria_Order,
  scale = 1,
  width = 17,
 height = 11,
 units = c("cm"),
  dpi = 600)
```



#consider the genus level
#https://www.geeksforgeeks.org/networkd3-package-in-r/
#https://adrientaudiere.github.io/MiscMetabar/articles/MiscMetabar.html
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
library("MiscMetabar")
library("phyloseq")
library("networkD3")
#summary_plot_pq(ps3_Bacteria)
#sankey_pq(gp, taxa = c(1:5))
#upset_pq(ps3_Bacteria, "Genotype", taxa = "Genus")

toptax = names(sort(taxa_sums(ps3_Bacteria), decreasing=TRUE))[1:25];toptax
ps.toptax_Bacteria = transform_sample_counts(ps3_Bacteria, function(x) {x/sum(x)*100});ps.toptax_Bacteria
ps.toptax_Bacteria <- prune_taxa(toptax, ps.toptax_Bacteria);ps.toptax_Bacteria
upset_bacteria_Genus=upset_pq(ps.toptax_Bacteria, "Genotype", taxa = "Genus");upset_bacteria_Genus

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Shared_upset_bacteria_Genus.tiff",     
 plot = upset_bacteria_Genus,
  scale = 1,
  width = 18,
 height = 11,
 units = c("cm"),
  dpi = 600)
```



##plot rarefaction curves  for the bacterial community
#code based on https://microsud.github.io/microbiomeutilities/articles/microbiomeutilities.html
```{r, echo=FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
library(microbiomeutilities)
# set seed
set.seed(1)
subsamples <- seq(0, 50000, by=10000)[-1]
#subsamples = c(10, 5000, 10000, 20000, 30000)
Rarefaction_Bacteria <- plot_alpha_rcurve(ps3_Bacteria, index="observed",
            subsamples = subsamples,
            lower.conf = 0.025, 
            upper.conf = 0.975,
            group="Genotype",
            label.color = "brown3",
            label.size = 3,
            label.min = TRUE);Rarefaction_Bacteria
Rarefaction_Bacteria <- Rarefaction_Bacteria + scale_color_manual(values = mycola) + 
  scale_fill_manual(values = mycola)+ 
  theme(axis.text=element_text(size=8), panel.grid=element_blank(),
        strip.background = element_rect(colour=NA,fill="grey33"),
        strip.text.x = element_text(face="bold"))+
  theme(text=element_text(size=14,  family="sans"))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, ncol=6))+
  theme(legend.position="right")+
  labs(tag = "")+    expand_limits(x = 0, y = 0); Rarefaction_Bacteria

#tiff("p_bact.tiff", units="in", width=12, height=5, res=300)
#ggsave("p_bact.tiff", plot = Rarefaction_Bacteria)

```

#Transformed to relative abundances 
#bar blot of the seed microbiome composition in the different cukltivars at the various locations for the top500 ASVs

```{r,echo=FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")

# Bar plots:
##Showing individual replicates. Top 500 taxa
toptax = names(sort(taxa_sums(ps3_Bacteria), decreasing=TRUE))[1:500];toptax
ps.toptax_Bacteria = transform_sample_counts(ps3_Bacteria, function(x) {x/sum(x)*100});ps.toptax_Bacteria
ps.toptax_Bacteria <- prune_taxa(toptax, ps.toptax_Bacteria);ps.toptax_Bacteria

#Class level
psmelt(ps.toptax_Bacteria)
gg1=ggplot(psmelt(ps.toptax_Bacteria), aes(x=reorder(Sample, Genotype), y=Abundance, fill=Class))+
  geom_bar(stat = "identity", position = "stack", width = 1) +
  #scale_fill_manual(values=c(ocean.dense(7), ocean.tempo(8)))+
  #scale_fill_manual(values=c("grey35","darkturquoise","#009E73",
   #      "Orange","#6A3D9A","#00AFBB","steelblue", "#cccda7", "#bfb77b", "7e9998", "#92B2B2", "#83bbbf", "#58a0c4","#8973b4", #"#56B4E9","#86486F","#446455", "bisque4"))+
  scale_fill_manual(values = mycolors)+
  xlab("") + ylab("Relative abundance (%)") +
  facet_wrap(~reorder(Genotype, Sample), scales = "free_x", nrow = 1) + 
  ggtitle("Class, top 500 ASVs") +
  theme_pubclean() + theme(legend.position="bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0)) +
theme(legend.text = element_text(face = c(rep("italic", 5), rep("plain", 5))))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, nrow = 3))+
  theme(text=element_text(size=8,  family="Arial"))+ labs(tag = "");gg1

library(ggpubr)
tiff("Composition_Genotype_bact.tiff", units="in", width=9, height=5, res=300)
Composition_Genotype_bact.tiff=gg1+plot_layout(nrow = 1);Composition_Genotype_bact.tiff
ggsave("Composition_Genotype_bact.tiff", plot = Composition_Genotype_bact.tiff)
dev.off()


#Order level
psmelt(ps.toptax_Bacteria)
gg2=ggplot(psmelt(ps.toptax_Bacteria), aes(x=reorder(Sample, Genotype), y=Abundance, fill=Order))+
  geom_bar(stat = "identity", position = "stack", width = 1) +
  #scale_fill_manual(values=c(ocean.dense(7), ocean.tempo(8)))+
  #scale_fill_manual(values=c("#25725d", "#439b5b", "#7bc369", "#abd69a", "#d8e8b0", "#cccda7", "#bfb77b", "#7e9998", "#92B2B2", "#83bbbf", "#61acc1", "#58a0c4", "#7087c3","#8973b4", "#512972"))+
  scale_fill_manual(values = mycola)+
  xlab("") + ylab("Relative abundance (%)") +
  facet_wrap(~reorder(Genotype, Sample), scales = "free_x", nrow = 1) + 
  ggtitle("Order, top 500 ASVs") +
  theme_pubclean() + theme(legend.position="bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0)) +
theme(legend.text = element_text(face = c(rep("italic", 5), rep("plain", 5))))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, nrow = 4))+
  theme(text=element_text(size=8,  family="Arial"))+ labs(tag = "");gg2

library(ggpubr)
tiff("Order_Composition_Genotype_bact.tiff", units="in", width=14, height=5, res=300)
Order_Composition_Genotype_bact.tiff=gg2+plot_layout(nrow = 1);Order_Composition_Genotype_bact.tiff
ggsave("Order_Composition_Genotype_bact.tiff", plot = Order_Composition_Genotype_bact.tiff)
dev.off()
```

#Average percentage composition class
```{r,echo=FALSE,warning=FALSE,message=FALSE}
 load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
#Convert to relative abundance or the computation of the micribial composition in the samples
#consider phylum
GPf_Phylum = tax_glom(ps3_Bacteria, "Phylum") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Phylum = data.frame(Phylum = tax_table(GPf_Phylum)[,"Phylum"], Mean = rowMeans(otu_table(GPf_Phylum)), row.names = NULL)
df_Phylum = df_Phylum[order(-df_Phylum$Mean),];df_Phylum
head(df_Phylum,20)

#consider Class
GPf_Class = tax_glom(ps3_Bacteria, "Class") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Class = data.frame(Class = tax_table(GPf_Class)[,"Class"], Mean = rowMeans(otu_table(GPf_Class)), row.names = NULL)
df_Class = df_Class[order(-df_Class$Mean),];df_Class
head(df_Class,20)

#consider Order
GPf_Order = tax_glom(ps3_Bacteria, "Order") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Order = data.frame(Order = tax_table(GPf_Order)[,"Order"], Mean = rowMeans(otu_table(GPf_Order)), row.names = NULL)
df_Order = df_Order[order(-df_Order$Mean),];df_Order
head(df_Order,20)

#consider Family
GPf_Family = tax_glom(ps3_Bacteria, "Family") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Family = data.frame(Family = tax_table(GPf_Family)[,"Family"], Mean = rowMeans(otu_table(GPf_Family)), row.names = NULL)
df_Family = df_Family[order(-df_Family$Mean),];df_Family
head(df_Family,20)

#consider Genus
GPf_Genus = tax_glom(ps3_Bacteria, "Genus") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Genus = data.frame(Genus = tax_table(GPf_Genus)[,"Genus"], Mean = rowMeans(otu_table(GPf_Genus)), row.names = NULL)
df_Genus = df_Genus[order(-df_Genus$Mean),];df_Genus
head(df_Genus,20)

```




#Beta diversity

```{r}
phyloseq_transform_css <- function(physeq, norm = TRUE, log = TRUE, ...){
  # require(metagenomeSeq)
  MGS <- phyloseq::phyloseq_to_metagenomeSeq(physeq)
  # get the normalized count matrix
  otu_norm <- metagenomeSeq::MRcounts(MGS, norm = norm, log = log, ...)
  # exportMat(datt.norm, file = "tmp.txt")    # norm = TRUE, log = TRUE
  ## save sample statistics (sample scaling factor, quantile value, number of identified features and library size):
  # exportStats(dattM.sf, file = "tmp_stats.txt")
  # Substitue raw abundance to the css-normalized data
  physeq.tr <- physeq
  phyloseq::otu_table(physeq.tr) <- phyloseq::otu_table(otu_norm, taxa_are_rows = T)
  return(physeq.tr)
}
```


```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
ps3_Bacteria_css = phyloseq_transform_css(ps3_Bacteria);ps3_Bacteria_css#########
#First get otu_table and transpose it:
dist.matrix_ps3_Bacteria_css <- t(data.frame(otu_table(ps3_Bacteria_css)))
sampledf_ps3_Bacteria_css <- data.frame(sample_data(ps3_Bacteria_css))
#Then use vegdist from vegan to generate a bray distance object:
phyloseq_bray_ps3_Bacteria_css <- vegdist(dist.matrix_ps3_Bacteria_css, method = "bray"); phyloseq_bray_ps3_Bacteria_css
sampleda <- data.frame(sample_data(ps3_Bacteria_css), check.names=FALSE);sampleda
sampleda <- sampleda[match(colnames(as.matrix(phyloseq_bray_ps3_Bacteria_css)),rownames(sampleda)),,drop=FALSE];sampleda
sampleda$Sample_type <- factor(sampleda$Sample_type)
adores <- adonis2(phyloseq_bray_ps3_Bacteria_css ~ Genotype, data=sampleda, permutation=999);adores#change from adonis to adonis2
#         Df SumOfSqs     R2      F Pr(>F)    
#Model     4   5.7450 0.7026 14.765  0.001 ***
#Residual 25   2.4318 0.2974                  
#Total    29   8.1768 1.0000                  
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
#############################
ps3_Bacteria_css = phyloseq_transform_css(ps3_Bacteria);ps3_Bacteria_css#########
#visualize
matrixdistance_ps3_Bacteria_css_PCoA <- ordinate(physeq = ps3_Bacteria_css, method = "PCoA", distance = "bray", trymax = 1000)
#Stress:     0.2069558 

pcoaplot1_bacteria.tiff=plot_ordination(physeq = ps3_Bacteria_css, ordination = matrixdistance_ps3_Bacteria_css_PCoA, color = "Genotype") + 
  theme(text = element_text(size = 12)) + scale_color_manual(values =mycola)+theme(legend.title=element_blank())+  geom_point(aes(color =Genotype ), size = 5)+
  scale_shape_manual(values=c(19,17, 15,18))+ theme_gray(base_size = 14)+ labs(tag = "C")+
  theme(legend.position="");pcoaplot1_bacteria.tiff
#tiff("pcoaplot1_bacteria.tiff", units="in", width=5, height=5, res=300)
#ggsave("pcoaplot1_bacteria.tiff", plot = pcoaplot1_bacteria.tiff)
#dev.off()

```


#Beta dispersion tests & plot
```{r}
dist=phyloseq_bray_ps3_Bacteria_css
dispr <- vegan::betadisper(dist, phyloseq::sample_data(ps3_Bacteria_css)$Genotype);dispr #calculate the dispersion

data_fram_cultivar <- data.frame(Distance_to_centroid=dispr$distances,Group=dispr$group);data_fram_cultivar
groups <- dispr$group

#stats
kruskal.test(Distance_to_centroid ~ Group, data = data_fram_cultivar)
#Kruskal-Wallis chi-squared = 8.6151, df = 4, p-value = 0.07148

#Decomposing group differences: If both PERMANOVA (tests for differences in centroids) and beta dispersion (tests for differences in variability) are significant, it suggests that groups differ not only in their average composition but also in their compositional spread. If only PERMANOVA is significant, the groups differ in their central tendency but have similar levels of variability. If only beta dispersion is significant, the groups have similar average compositions but differ in their
```


#alpha diversity
```{r}
 load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")

Bacteria_subsamp <- rarefy_even_depth(ps3_Bacteria, sample.size=10000  , rngseed=5163, replace=FALSE, trimOTUs=TRUE);Bacteria_subsamp#[ 1128 samples by 20 sample variables ]
Bacterial_richness = estimate_richness(Bacteria_subsamp , measures=c("Observed","Shannon"))
Bacterial_richness_table <- as.data.frame(as.matrix(Bacterial_richness))
write.csv(Bacterial_richness_table,file = "Bacterial_richness_table.csv")
Bacterial_richness <- read.table("Bacterial_richness_table.txt", header=TRUE); Bacterial_richness
```


#Bacterial diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Bacterial_richness <- read.table("Bacterial_richness_table.txt", header=TRUE); Bacterial_richness

plot_bacterial_rich=ggboxplot(Bacterial_richness, x = "Genotype", y = "Observed",fill  = "Genotype", ylab = "Observed ASVs", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="bottom") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_bacterial_rich

tiff("plot_bacterial_rich.tiff", units="in", width=12, height=3, res=300)
plot_bacterial_rich.tiff=plot_bacterial_rich+plot_layout(ncol =1 );plot_bacterial_rich.tiff
ggsave("plot_bacterial_rich.tiff", plot = plot_bacterial_rich.tiff)
dev.off()


tiff("Bacterial_rich.tiff", units="in", width=2, height=3, res=300)
Bacterial_rich.tiff=plot_bacterial_rich+plot_layout(ncol =1 );Bacterial_rich.tiff
ggsave("Bacterial_rich.tiff", plot = Bacterial_rich.tiff)
dev.off()



plot_bacterial_Shannon=ggboxplot(Bacterial_richness, x = "Genotype", y = "Shannon",fill  = "Genotype", ylab = "Shannon index", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_bacterial_Shannon


tiff("Bacterial_Shannon.tiff", units="in", width=2, height=3, res=300)
Bacterial_Shannon.tiff=plot_bacterial_Shannon+plot_layout(ncol =1 );Bacterial_Shannon.tiff
ggsave("Bacterial_Shannon.tiff", plot = Bacterial_Shannon.tiff)
dev.off()
```



#Bacterial diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Bacterial_richness <- read.table("Bacterial_richness_table.txt", header=TRUE); Bacterial_richness
attach(Bacterial_richness)

#Consider microbial richness
#Normality
shapiro.test(Observed)
#W = 0.78301, p-value = 4.204e-05

#statistics measures for the different bacterial indices
kruskal.test(Observed ~ Genotype, data = Bacterial_richness)

#data:  Observed by Genotype
#Kruskal-Wallis chi-squared = 26.151, df = 4, p-value = 2.95e-05

#Post-hoc (Dunn's tests)


Bacterial_Dunn_observed_genotype=dunn_test(Bacterial_richness, Observed ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Bacterial_Dunn_observed_genotype
write_csv(Bacterial_Dunn_observed_genotype, "Bacterial_Dunn_observed_genotype.csv")
Bacterial_Dunn_observed_genotype <- dunnTest(Observed, Genotype, method = "bonferroni");Bacterial_Dunn_observed_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Bacterial_Dunn_observed_genotype = cldList(P.adj ~ Comparison, data=Bacterial_Dunn_observed_genotype$res);CLD_Bacterial_Dunn_observed_genotype 
```


#Bacterial diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Bacterial_richness <- read.table("Bacterial_richness_table.txt", header=TRUE); Bacterial_richness
attach(Bacterial_richness)

#Consider microbial Shannon
#Normality
shapiro.test(Shannon)
#W = 0.85632, p-value = 0.001029

#statistics measures for the different bacterial indices
kruskal.test(Shannon ~ Genotype, data = Bacterial_richness)

#data:  Shannon by Genotype
#Kruskal-Wallis chi-squared = 26.184, df = 4, p-value = 2.905e-05

#Post-hoc (Dunn's tests)

Bacterial_Dunn_Shannon_genotype=dunn_test(Bacterial_richness, Shannon ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Bacterial_Dunn_Shannon_genotype
write_csv(Bacterial_Dunn_Shannon_genotype, "Bacterial_Dunn_Shannon_genotype.csv")
Bacterial_Dunn_Shannon_genotype <- dunnTest(Shannon, Genotype, method = "bonferroni");Bacterial_Dunn_Shannon_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Bacterial_Dunn_Shannon_genotype = cldList(P.adj ~ Comparison, data=Bacterial_Dunn_Shannon_genotype$res);CLD_Bacterial_Dunn_Shannon_genotype 

```


#Bacterial abundance

```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Bacterial_Abundance <- read.table("Bacteria_abundance.txt", header=TRUE, sep="\t"); Bacterial_Abundance

attach(Bacterial_Abundance)
variable.names(Bacterial_Abundance)
Bacterial_Abundance$log10
head(Bacterial_Abundance)
str(Bacterial_Abundance)

plot_bacterial_Abundance=ggboxplot(Bacterial_Abundance, x = "Genotype", y = "log10",fill  = "Genotype", ylab = "Log(10) gene copies/gram", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_bacterial_Abundance


tiff("Bacterial_abundance.tiff", units="in", width=2, height=3, res=300)
Bacterial_abundance.tiff=plot_bacterial_Abundance+plot_layout(ncol =1 );Bacterial_abundance.tiff
ggsave("Bacterial_abundance.tiff", plot = Bacterial_abundance.tiff)
dev.off()



#statistics
#Normality
shapiro.test(log10)
#W = 0.85632, p-value = 0.001029

#statistics measures for the different bacterial indices
kruskal.test(log10 ~ Genotype, data = Bacterial_Abundance)

#data:  log10 by Genotype
#Kruskal-Wallis chi-squared = 26.184, df = 4, p-value = 2.905e-05

#Post-hoc (Dunn's tests)


Bacterial_Dunn_log10_genotype=dunn_test(Bacterial_Abundance, log10 ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Bacterial_Dunn_log10_genotype
write_csv(Bacterial_Dunn_log10_genotype, "Bacterial_Dunn_log10_genotype.csv")
Bacterial_Dunn_log10_genotype <- dunnTest(log10, Genotype, method = "bonferroni");Bacterial_Dunn_log10_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Bacterial_Dunn_log10_genotype = cldList(P.adj ~ Comparison, data=Bacterial_Dunn_log10_genotype$res);CLD_Bacterial_Dunn_log10_genotype 

```


#Differentially abundant bacterial taxa

```{r}
#https://rpubs.com/kjojess/1013527
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
#Install corncob using:
#remotes::install_github("statdivlab/corncob")
library(corncob)
library(lefser)
require(dplyr)
require(tibble)
library(qiime2R)
library(circlize)
library(ggplot2)
library(microbiomeMarker)
library(RColorBrewer)
#library(ggthemr)
library(plyr)
library(phyloseq)
library(tidyverse)
library(ComplexHeatmap)
library(cowplot)
library(phyloseq)
library(microbiomeMarker)
#Load the data
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")
set.seed(1)
#sample_data(ps3_Bacteria)

setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

lefse_cc_path_Genus <- run_lefse(
    ps3_Bacteria,
    wilcoxon_cutoff = 0.05,
    group = "Genotype",
    kw_cutoff = 0.05,
    multigrp_strat = TRUE,
    lda_cutoff =4,
    taxa_rank="Genus"
)
lefse_cc_path_Genus
sig_marker_lefse<-data.frame(marker_table(lefse_cc_path_Genus))

plot_abundance(lefse_cc_path_Genus, group="Genotype")
Lefse_bacteria=plot_ef_bar(lefse_cc_path_Genus);Lefse_bacteria

```


#Random forest classifiers of plant genpotype (using the seed microbiome)
#Load packages 
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
# Define prevalence of each taxa; # (in how many samples did each taxa appear at least once)
library(doParallel);packageVersion("doParallel")#‘1.0.17’
library(foreach);packageVersion("foreach")#‘1.5.2’
library(tidyverse);packageVersion("tidyverse")#‘2.0.0’
library(ggplot2);packageVersion("ggplot2")#‘3.5.0’
library(vegan);packageVersion("vegan")#‘2.6.4’
library(reshape2);packageVersion("reshape2")#‘1.4.4’
library(phyloseq);packageVersion("phyloseq")#‘1.46.0’
library(randomForest);packageVersion("randomForest")
library(doBy);packageVersion("doBy")#‘4.7.1.1’
library(plotrix);packageVersion("plotrix")#‘3.8.4’
library(ggpubr);packageVersion("ggpubr")#‘0.6.0’

Bacteria_otu <- read.table("Bacteria_Table-rf.txt", header=TRUE,row.names="OTUID");Bacteria_otu
Bacteria_otu <- otu_table(Bacteria_otu, taxa_are_rows = TRUE);Bacteria_otu 

## Imported taxonomy
#Bacteria_tax <- read.csv("Taxonomy.csv",header=TRUE,row.names="OTUID");Bacteria_tax
Bacteria_tax <- read.table("Bacteria_Taxonomy-rf.txt", header=TRUE, sep="\t",row.names="OTUID");Bacteria_tax
Bacteria_tax <- as.matrix(Bacteria_tax);Bacteria_tax
Bacteria_tax <- tax_table(Bacteria_tax);Bacteria_tax

## Imported metadata file
#metadata <- read.table("Metadata.txt", header=TRUE, sep="\t",row.names="SampleID");metadata
Bacteria_metadata<- read.csv("Bacteria_metadata.csv",header=TRUE,row.names="SampleID");Bacteria_metadata

#rownames(Bacteria_metadata) <- Bacteria_metadata[,1]
Bacteria_metadata <- sample_data(Bacteria_metadata);Bacteria_metadata
#tree
#tree <- read.tree(file = "tree.nwk");tree
#tree <- phy_tree(tree)

all(rownames(Bacteria_metadata) %in% colnames(Bacteria_otu))
#Convert to phyloseq object
Bacteria <- merge_phyloseq(Bacteria_otu,Bacteria_tax, Bacteria_metadata);Bacteria#[ 16460 taxa and 1197 samples ]
Bacteria = subset_taxa(Bacteria, Kingdom=="Bacteria");Bacteria# [ 6448 taxa and 84 samples ]
Bacteria = subset_taxa(Bacteria, Phylum!="CyanoBacteria");Bacteria#  [ 15588 taxa and 1197 samples ]
Bacteria = subset_taxa(Bacteria, Order!="Chloroplast");Bacteria#  [ 15588 taxa and 1197 samples ]
Bacteria = subset_taxa(Bacteria, Family!="Mitochondria");Bacteria# [ 15588 taxa and 1197 samples ]
Bacteria <- prune_taxa(taxa_sums(Bacteria)>0, Bacteria);Bacteria# [ 15588 taxa and 1197 samples ]
#OTU_table_Bacteria = as(otu_table(Bacteria), "matrix")
#########################################################################################################
#Summarize 
microbiome::summarize_phyloseq(Bacteria)
#Number of singletons = 68"
#Sparsity = 0.995179763625127"
#Average number of reads = 22794.9197994987"
#Total number of reads = 27285519"
#Max. number of reads = 225892"
#Min. number of reads = 68"
sums_Bacteria <- sample_sums(Bacteria);sums_Bacteria
sums_Bacteria <- data.frame(sums_Bacteria);sums_Bacteria
sums_Bacteria <- as.matrix(sums_Bacteria);sums_Bacteria
colSums(sums_Bacteria) #  27281011      
max(sums_Bacteria) #225892
min(sums_Bacteria) # 68
sums_Bacteria
#https://david-barnett.github.io/microViz/articles/web-only/tax-fixing.html #fixing the table
#https://rpubs.com/bioguo/680490 (Random forest classifications)
library(devtools);packageVersion("devtools")# ‘2.4.5’# Load the devtools package
#install_github("guokai8/microbial") # Install the package (Microbial package)
library(microbial);packageVersion("microbial")#‘0.0.22’####
ps_RandomForest_bacteria=Bacteria
save(ps_RandomForest_bacteria,
file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps_RandomForest_bacteria.RData")



#Bacterial claasifier
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps_RandomForest_bacteria.RData")

prunescale = 0.05

# Prune out rare OTUs by mean relative abundance set by prunescale
tax.mean.ps_RandomForest_bacteria <- taxa_sums(ps_RandomForest_bacteria)/nsamples(ps_RandomForest_bacteria) # average number of reads for each otu
sites.prune.ps_RandomForest_bacteria <- prune_taxa(tax.mean.ps_RandomForest_bacteria > prunescale, ps_RandomForest_bacteria)

#Next, we are going to format a dataframe of predictors (OTUs) and responses (States)
# Make a dataframe of training data with OTUs as column and samples as rows
predictors.ps_RandomForest_bacteria<- t(otu_table(sites.prune.ps_RandomForest_bacteria));predictors.ps_RandomForest_bacteria
dim(predictors.ps_RandomForest_bacteria)
# 1197 1655

# Make one column for our outcome/response variable 
response.ps_RandomForest_bacteria <- as.factor(sample_data(sites.prune.ps_RandomForest_bacteria)$Genotype);response.ps_RandomForest_bacteria
# Combine them into 1 data frame
rf.data.ps_RandomForest_bacteria <- data.frame(response.ps_RandomForest_bacteria, predictors.ps_RandomForest_bacteria);rf.data.ps_RandomForest_bacteria
str(rf.data.ps_RandomForest_bacteria)
# Set a random seed
set.seed(5000)
# My computer has 8 processor cores - adjust parameter of the following command as needed/desired
registerDoParallel(3)

# %dopar% implemented with foreach implements each loop call as an independent function call
# and then parallelizes all the function calls across the number of processers set above with
# registerDoParallel(). First, we need to set up a dataframe to store each bootstrap RF result.
boot.imp.s <- data.frame(predictors=NULL, Try=NULL, MeanDecreaseGini=NULL)
boot.imp.s <- foreach (Try=1:100, .combine=rbind) %dopar% {
  
  ps_RandomForest_bacteria.classify <- randomForest(response.ps_RandomForest_bacteria ~., data = rf.data.ps_RandomForest_bacteria, ntree = 500, proximity=T, mtry=200)
  print(ps_RandomForest_bacteria.classify)
  
  # Make a data frame with predictor names and their importance
  imp.s <- importance(ps_RandomForest_bacteria.classify)
  imp.s <- data.frame(predictors = rownames(imp.s), imp.s)
  
  cbind(imp.s, data.frame(Try=Try)) #)
  
}

# Same (and independent) as above but this time we are storing results from the
# out-of-bag error rates, so we car report the correct-incorrect classification
# rate for each group, and what types of misclassifications were made.
boot.oob <- NULL
boot.oob <- foreach (Try=1:100, .combine=rbind) %dopar% {
  
  ps_RandomForest_bacteria.classify <- randomForest(response.ps_RandomForest_bacteria ~., data = rf.data.ps_RandomForest_bacteria, ntree = 500, proximity=T, mtry=200)
  print(ps_RandomForest_bacteria.classify)
  
  # Make a data frame with predictor names and their importance
  imp.s <- importance(ps_RandomForest_bacteria.classify)
  imp.s <- data.frame(predictors = rownames(imp.s), imp.s)
  
  c(boot.oob, ps_RandomForest_bacteria.classify$err.rate[500,1])
}

#Display and visualize results
# Look at the results
mean(boot.oob)#0.1007352
sd(boot.oob)#0.003860804
range(boot.oob)# 0.09106099 0.11027569
# Create a dataframe with summary statistics from our bootstrap runs
summary.boot.s <- summaryBy(MeanDecreaseGini ~ predictors, data= boot.imp.s, FUN = c(mean, sd, std.error));summary.boot.s
write.csv(summary.boot.s,"summary.boot.s.csv")
# Look at it
head(summary.boot.s)
str(summary.boot.s)
# Order the predictor levels by importance
imp.s.sort <- arrange(summary.boot.s, desc(MeanDecreaseGini.mean));imp.s.sort
imp.s.sort$predictors <- factor(imp.s.sort$predictors, levels = imp.s.sort$predictors)
# Select the top 50 predictors
imp.s.5 <- imp.s.sort[1:20, ];imp.s.5

# What are those OTUs?
# otunames.s = otunames of predictors
# r.s = A logical containing which rows from phyloseq object were recovered as top 40 predictors
otunames.s <- imp.s.5$predictors;otunames.s
r.s <- rownames(tax_table(ps_RandomForest_bacteria)) %in% otunames.s;r.s

# Use r.s to pull out the taxonomic information for those OTUS in otunames.s
t.table <- as.data.frame(tax_table(ps_RandomForest_bacteria)[r.s, ])
write.csv(t.table, "RF.ps_RandomForest_bacteria.t.table.csv", quote=F,row.names=T)


#VISUALIZATION HERE
# Reformat the taxonomic names and paste them to OTU names for axis labels
# (highest determined taxonomic rate, get rid of special characters, etc.)
axislabels <- with(t.table, paste(rownames(t.table), gsub("[sgfocp]__","",Genus,perl=T) %>% gsub("_"," ",.,perl=T) %>% gsub("[\\s]?(unclassified|ge)[\\s]?","",.,perl=T)))
names(axislabels) <- rownames(t.table)
i <- grep("(uncultured)|([01234567890-]$)+", axislabels, perl=T)
axislabels[i]<-paste(names(axislabels)[i], t.table$Family[i])

# Now we sum the abundance of each OTU by state so we can order the
# RF predictor OTUs by which state they were most abundant in, followed
# by their importance value. This makes the heatmap easier to interpret.
sums<-summaryBy(. ~ response.ps_RandomForest_bacteria, data = rf.data.ps_RandomForest_bacteria[, c('response.ps_RandomForest_bacteria', as.character(otunames.s))], FUN = sum, keep.names = T);sums
state.organize <- sapply(as.character(otunames.s), FUN = function(x) which.max(sums[,x]));state.organize
sums$response.ps_RandomForest_bacteria
rf.data.ps_RandomForest_bacteria$response.ps_RandomForest_bacteria
# hm is a dataframe of the abundance of the predictors, standardized and log transformed
hm <- rf.data.ps_RandomForest_bacteria[,as.character(otunames.s)] %>% decostand(method='max', MARGIN=2) %>% decostand(method='log');hm
hm$sample <- gsub('_',' ',sample_data(ps_RandomForest_bacteria)[rownames(hm),'Genotype']$Genotype)

#hm.melt <- melt(hm[c(order(state.organize), 6)]);hm.melt
hm.melt <-  as.data.table(hm[c(order(state.organize), 21)]);hm.melt
hm.melt <- melt(hm.melt, id.vars = "sample")
names(hm.melt) <- c("Tree","Taxon","StandardizedAbundance")

# Turn predictors into a factor (re-ordered the same as the hm.melt object so they match up)
imp.s.5$predictors <- factor(imp.s.5$predictors, levels(imp.s.5$predictors)[order(state.organize)])


# Creat the heatmap with the taxon names and OTUs
plot1_bacteria <- ggplot(hm.melt, aes(x=Tree, y=Taxon, fill=StandardizedAbundance)) + 
  geom_tile() + scale_fill_gradient(low="white", high="gray33") + 
  theme(axis.text.x = element_text(angle=90, size=11), legend.position = "none", axis.text.y = element_text(hjust=0))  + 
  scale_y_discrete(labels=axislabels)+theme(axis.title.x = element_text(face="bold", colour="black", size=11),
                                            axis.title.y = element_text(face="bold", colour="black", size=11),
                                            axis.text.x = element_text(face="bold", colour="black", size=11),
                                            axis.text.y  = element_text(face="bold", colour="black", size=11))+ labs(tag = "",size=11);plot1_bacteria

#Creat the bargraph
plot2_bacteria <- ggplot(imp.s.5, aes(x = predictors, y = MeanDecreaseGini.mean)) +
  geom_bar(stat = "identity", fill = "bisque4") +
  ylab("Bootstrap Mean \u00b1 SD Gini Index") +
  theme_bw() +
  geom_errorbar( aes(x=predictors, ymin=MeanDecreaseGini.mean-MeanDecreaseGini.sd, ymax=MeanDecreaseGini.mean+MeanDecreaseGini.sd), width=0.75, alpha=0.9, size=0.5) +
  coord_flip() + theme(axis.title.x = element_text(face="bold", colour="black", size=11),
                       axis.title.y = element_text(face="bold", colour="black", size=11),
                       axis.text.x = element_text(face="bold", colour="black", size=11),
                       axis.text.y  = element_text(face="bold", colour="black", size=11))+ labs(tag = "",size=11);plot2_bacteria
g2_bacteria <- ggplotGrob(plot2_bacteria);g2_bacteria

library(ggpubr)
library(patchwork)

#Put them together

tiff("RF1_bacteria.tiff", units="in", width=15, height=6, res=300)
RF1_bacteria.tiff=plot1_bacteria+plot2_bacteria+plot_layout(nrow=1);RF1_bacteria.tiff
ggsave("RF1_bacteria.tiff", plot = RF1_bacteria.tiff)
dev.off()
```


************************************************************************************************
#Fungal community


#Load data (ASV table, taxonomy & tree, & metadata_Fungi)
```{r,echo = F}
##Imported feature table
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

asv_tab_Fungi <- read.table("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Fungi_Table.txt", 
              header=TRUE,
              row.names="OTUID");asv_tab_Fungi
asv_tab_Fungi <- otu_table(asv_tab_Fungi, taxa_are_rows = TRUE);asv_tab_Fungi 

## Imported taxonomy table
tax_tab_Fungi <- read.table("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Fungi_Taxonomy.txt", 
              header=TRUE,
              sep="\t",
              row.names="OTUID");tax_tab_Fungi
tax_tab_Fungi <- as.matrix(tax_tab_Fungi);tax_tab_Fungi
tax_tab_Fungi <- tax_table(tax_tab_Fungi);tax_tab_Fungi

## Imported metadata_Fungi file
metadata_Fungi <- read.csv("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Fungi_Metadata.csv",
              header=TRUE,
              row.names="SampleID");metadata_Fungi
#rownames(metadata_Fungi) <- metadata_Fungi[,1]
metadata_Fungi <- sample_data(metadata_Fungi);metadata_Fungi

#import reference sequences
## Import reference sequences
# rs_table <- Biostrings::readDNAStringSet(
#   "outputs/qiime2/dna-sequences.fasta",
#   format = "fasta",
#   nrec = -1L,
#   skip = 0L,
#   seek.first.rec = F,
#   use.names = T,
#   with.qualities = F)

#proof that the data are of equal dimensions
all(rownames(metadata_Fungi) %in% colnames(asv_tab_Fungi))

#Convert to phyloseq object
ps_Fungi <- merge_phyloseq(
  asv_tab_Fungi,
  tax_tab_Fungi, 
  metadata_Fungi);ps_Fungi#[ 16460 taxa & 1197 samples ]
summarize_phyloseq(ps_Fungi)
ps0_Fungi <- ps_Fungi # 16460 taxa & 1197 samples
save(ps0_Fungi,
file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps0_Fungi.RData")

```


#clean the data
```{r,echo = FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps0_Fungi.RData")

ps_clean_Fungi = subset_taxa(ps0_Fungi, Kingdom=="Fungi");ps_clean_Fungi# [ 6448 taxa & 84 samples ]
#ps_clean_Fungi = subset_taxa(ps_clean_Fungi, Phylum!="CyanoFungi");ps_clean_Fungi#  [ 15588 taxa & 1197 samples ]
#ps_clean_Fungi = subset_taxa(ps_clean_Fungi, Order!="Chloroplast");ps_clean_Fungi#  [ 15588 taxa & 1197 samples ]
#ps_clean_Fungi = subset_taxa(ps_clean_Fungi, Family!="Mitochondria");ps_clean_Fungi# [ 15588 taxa & 1197 samples ]
ps_clean_Fungi <- prune_taxa(taxa_sums(ps_clean_Fungi)>0, ps_clean_Fungi);ps_clean_Fungi# [ 15588 taxa & 1197 samples ]

ps1_Fungi <- ps_clean_Fungi

sort(sample_sums(ps1_Fungi))
summary(sample_sums(ps1_Fungi)) # mean reads/sample = 22791
summarize_phyloseq(ps1_Fungi) # Average reads = 22791, Total reads = 27281011, Singletons = 38
taxa_names(ps1_Fungi)[1:5]
sample_data(ps1_Fungi)$total_reads <- sample_sums(ps1_Fungi)

save(ps1_Fungi, file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps1_Fungi.RData")

K <- as.data.frame(table(tax_table(ps1_Fungi)[,"Kingdom"]));K
P <- as.data.frame(table(tax_table(ps1_Fungi)[,"Phylum"]));P # 9 phyla
C <- as.data.frame(table(tax_table(ps1_Fungi)[,"Class"]));C # 18 classes
O <- as.data.frame(table(tax_table(ps1_Fungi)[,"Order"]));O # 42 orders
Fa <- as.data.frame(table(tax_table(ps1_Fungi)[,"Family"]));Fa # 70 families
G <- as.data.frame(table(tax_table(ps1_Fungi)[,"Genus"]));G # 104 genera
#S <- as.data.frame(table(tax_table(ps1_Fungi)[,"Species"]));S # 1264 species

microbiome::summarize_phyloseq(ps1_Fungi)
#Number of singletons = 68"
#Sparsity = 0.995179763625127"
#Average number of reads = 22794.9197994987"
# Total number of reads = 27281011"
#Max. number of reads = 225892"
#Min. number of reads = 68"
taxonomy <- as.vector(phyloseq::tax_table(ps1_Fungi))
sums_ps1_Fungi <- sample_sums(ps1_Fungi);sums_ps1_Fungi
sums_ps1_Fungi <- data.frame(sums_ps1_Fungi);sums_ps1_Fungi
sums_ps1_Fungi <- as.matrix(sums_ps1_Fungi);sums_ps1_Fungi
colSums(sums_ps1_Fungi) #  27281011      
max(sums_ps1_Fungi) #225892
min(sums_ps1_Fungi) # 68
sums_ps1_Fungi
```


#Filter data at 0.025 prevalence

#Obtained from: https://f1000research.com/articles/5-1492/v1;#One of the reasons to filter by prevalence is to avoid spending #much time analyzing taxa that were only rarely seen. This also turns out to be a useful filter of noise (taxa that are #actually just artifacts of the data collection process)
# Define prevalence of each taxa; # (in how many samples did each taxa appear at least once)

```{r,echo = FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps1_Fungi.RData")

prev0 = apply(X = otu_table(ps1_Fungi),
              MARGIN = ifelse(taxa_are_rows(ps1_Fungi), yes = 1, no = 2),
              FUN = function(x){sum(x > 0)})

prevdf = data.frame(Prevalence = prev0,
                    TotalAbundance = taxa_sums(ps1_Fungi),
                    tax_table(ps1_Fungi))# the most prevalent ASVs present therein the composition

#only for reference, in case we want to filter Phyla that appear less than X times
keepPhyla = table(prevdf$Phylum)[(table(prevdf$Phylum) > 0)]

prevdf1 = subset(prevdf, Phylum %in% names(keepPhyla))

# Define prevalence threshold as 0.25% of total samples# 
prevalenceThreshold = 0.0025 * nsamples(ps1_Fungi)
prevalenceThreshold

# Execute prevalence filter, using `prune_taxa()` function
ps2_Fungi_Fungi = prune_taxa((prev0 > prevalenceThreshold), ps1_Fungi);ps2_Fungi_Fungi

# Filter entries with unidentified Phylum.
ps3_Fungi = subset_taxa(ps2_Fungi_Fungi, Phylum %in% names(keepPhyla));ps3_Fungi

#Plot prevalence of different phyla
#https://github.com/joey711/phyloseq/issues/1474

ggplot(prevdf1, aes(TotalAbundance, Prevalence, color = Phylum)) +
  geom_hline(yintercept = prevalenceThreshold, alpha = 0.5, linetype = 2) +
  geom_point(size = 2, alpha = 0.65) +
  scale_y_log10() + scale_x_log10() +
  ggtitle(paste0("Prevalence filtering, using a threshold of ", prevalenceThreshold))+
  xlab("Total Abundance") +
  scale_color_manual(values=ocean.balance(38))+
  theme_bw()+
  facet_wrap(~Phylum, nrow = 4)

print(paste0("Initial #ASVs:", ntaxa(ps1_Fungi), ". #ASVs after prevalence filtering:", ntaxa(ps3_Fungi), ". #ASVs removed: ", ntaxa(ps1_Fungi)-ntaxa(ps3_Fungi)))
```



# Save the Fungi object
```{r}
save(ps3_Fungi, file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
```




## Ab-Prev prep
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
library(grid)   # make sure grid::margin is active
meta <- microbiome::meta(ps3_Fungi) # x_bac
asv <- otu_table(ps3_Fungi) # x_bac
taxo <- rownames_to_column(as.data.frame(tax_table(ps3_Fungi)),
                                        "ASV") # x_bac
sample_names(ps3_Fungi)
# presence-absence data
asv_PA <- 1 * ((asv > 0) == 1)

# occupancy calculation
asv_prev <- rowSums(asv_PA) / ncol(asv_PA)

# relative abundance
asv_rel <- apply(vegan::decostand(asv, 
                                  method = "total", 
                                  MARGIN = 2), 1, mean)

# combine occupancy and RA of each ASV in a table
asv_prev_rel <- rownames_to_column(as.data.frame(cbind(asv_prev, 
                                                       asv_rel)),
                                   "ASV")

asv_prev_rel_taxo <- merge(taxo, 
                           asv_prev_rel, 
                           by = "ASV")

# merge with metadata
asv_tdf <- rownames_to_column(as.data.frame(t(asv)),
                           "replicates")
library(dplyr)
meta <- meta <- meta %>% rename_with(~ "replicates", 1)
colnames(meta)

colnames(asv_tdf)
asv_tdf_meta <- merge(meta,
                      asv_tdf, 
                      by = "replicates")

# keep only the one feature column for collaps3_Fungiing table
asv_tdf_meta <- asv_tdf_meta[-c(1, 3:29)]

asv_gen <- asv_tdf_meta %>% 
    group_by(Genotype) %>%
  summarise_each(list(sum))

# make a ASV P/A table
asv_gen_PA <- 1 * ((asv_gen > 0) == 1)

# merge prevalence, RA and taxa with the nr of obs per genotype
asv_gen_obs <- data.frame(colSums(asv_gen_PA)) %>%
  na.omit() %>% 
  rownames_to_column("ASV")
names(asv_gen_obs)[names(asv_gen_obs) == "colSums.asv_gen_PA."] <- 'Obs'

asv_prev_rel_taxo_gen_obs <- merge(asv_prev_rel_taxo, 
                                   asv_gen_obs, 
                                   by = "ASV")
```




## Ab-Prev Plot 
```{r}
counts <- asv_prev_rel_taxo_gen_obs %>% dplyr::count(Kingdom = "Fungi")

p_core_flex1_all_Fungi <- 
  ggplot(asv_prev_rel_taxo_gen_obs, 
         aes(x = asv_rel, y = asv_prev, size = Obs)) +
  geom_point(
    color = "darkgrey",
    alpha = 0.5) + 
  ggrepel::geom_text_repel(data = subset(asv_prev_rel_taxo_gen_obs, asv_prev > 0.5),
                           max.overlaps3_Fungi = 20,
                           aes(asv_rel,
                               asv_prev,
                               label = Genus),
                           size = 3,
                           show.legend = F) +
  geom_text(data = counts,
            aes(x = 0.00000001, y = 0.1, 
                label = paste0("Total ASVs\n", n),  
                fontface = "bold"), size = 4) +
  geom_hline(yintercept = 0.9, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.7, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.5, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.25, linetype = "dotted", color = "red") +
  geom_vline(xintercept = 0.00005, linetype = "dotted", color = "red", show.legend = T) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.line = element_blank(),
    legend.text = element_text(size = 12),
    legend.position = c(0.35, 0.95),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "transparent", color = "transparent"),
    legend.margin = margin(0, 0, 0, 0),
    panel.spacing = unit(0.1, "lines")) +
  scale_x_log10(breaks = c(0.1, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001), 
                labels = c(10, 1, 0.1, 0.01, 0.001, 0.0001, 0.00001)) +
  scale_y_continuous(limits = c(-0.01, 1.01),
                     breaks = c(0,0.25, 0.5, 0.7, 0.9),
                     labels = c(0,25, 50, 70, 90)) +
  labs(x = "Relative abundance [%]", y = "Prevalence [%]", size = "nº cultivars"); p_core_flex1_all_Fungi

 ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ab-prev-all_Fungi.tiff",     
  plot = p_core_flex1_all_Fungi,
   scale = 1,
 width = 25,
  height = 20,
  units = c("cm"),
 dpi = 600)
```




#comsider order level
#https://www.geeksforgeeks.org/networkd3-package-in-r/
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
library("MiscMetabar")
library("phyloseq")
library("networkD3")
#summary_plot_pq(ps3_Fungi)
#sankey_pq(gp, taxa = c(1:5))
#upset_pq(ps3_Fungi, "Genotype", taxa = "Genus")

toptax = names(sort(taxa_sums(ps3_Fungi), decreasing=TRUE))[1:100];toptax
ps.toptax_Fungi = transform_sample_counts(ps3_Fungi, function(x) {x/sum(x)*100});ps.toptax_Fungi
ps.toptax_Fungi <- prune_taxa(toptax, ps.toptax_Fungi);ps.toptax_Fungi
upset_Fungi_Order=upset_pq(ps.toptax_Fungi, "Genotype", taxa = "Order");upset_Fungi_Order

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Shared_upset_Fungi_Order.tiff",     
  plot = upset_Fungi_Order,
  scale = 1,
  width = 20,
 height = 11,
 units = c("cm"),
  dpi = 600)
```



#consider the genus level
#https://www.geeksforgeeks.org/networkd3-package-in-r/
#https://adrientaudiere.github.io/MiscMetabar/articles/MiscMetabar.html
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
library("MiscMetabar")
library("phyloseq")
library("networkD3")
#summary_plot_pq(ps3_Fungi)
#sankey_pq(gp, taxa = c(1:5))
#upset_pq(ps3_Fungi, "Genotype", taxa = "Genus")

toptax = names(sort(taxa_sums(ps3_Fungi), decreasing=TRUE))[1:25];toptax
ps.toptax_Fungi = transform_sample_counts(ps3_Fungi, function(x) {x/sum(x)*100});ps.toptax_Fungi
ps.toptax_Fungi <- prune_taxa(toptax, ps.toptax_Fungi);ps.toptax_Fungi
upset_Fungi_Genus=upset_pq(ps.toptax_Fungi, "Genotype", taxa = "Genus");upset_Fungi_Genus

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Shared_upset_Fungi_Genus.tiff",     
 plot = upset_Fungi_Genus,
  scale = 1,
  width = 18,
 height = 11,
 units = c("cm"),
  dpi = 600)
```




##plot rarefaction curves  for the Fungi community
#code based on https://microsud.github.io/microbiomeutilities/articles/microbiomeutilities.html
```{r, echo=FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
library(microbiomeutilities)
# set seed
set.seed(1)
subsamples <- seq(0, 50000, by=5500)[-1]
#subsamples = c(10, 5000, 10000, 20000, 30000)

Rareaction_Fungi <- plot_alpha_rcurve(ps3_Fungi, index="observed",
            subsamples = subsamples,
            lower.conf = 0.025, 
            upper.conf = 0.975,
            group="Genotype",
            label.color = "brown3",
            label.size = 3,
            label.min = TRUE);Rareaction_Fungi

Rareaction_Fungi <- Rareaction_Fungi + scale_color_manual(values = mycola) + 
  scale_fill_manual(values = mycola)+ 
  theme(axis.text=element_text(size=8), panel.grid=element_blank(),
        strip.background = element_rect(colour=NA,fill="grey33"),
        strip.text.x = element_text(face="bold"))+
  theme(text=element_text(size=14,  family="sans"))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, ncol=6))+
  theme(legend.position="none")+
  labs(tag = "")+    expand_limits(x = 0, y = 0); Rareaction_Fungi

tiff("p_Fungi.tiff", units="in", width=12, height=5, res=300)
ggsave("p_Fungi.tiff", plot = Rareaction_Fungi)

```

#Transformed to relative abundances 
#bar blot of the seed microbiome composition in the different cukltivars at the various locations for the top500 ASVs

```{r,echo=FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")

# Bar plots:
##Showing individual replicates. Top 500 taxa
toptax = names(sort(taxa_sums(ps3_Fungi), decreasing=TRUE))[1:500];toptax
ps.toptax_Fungi = transform_sample_counts(ps3_Fungi, function(x) {x/sum(x)*100});ps.toptax_Fungi
ps.toptax_Fungi <- prune_taxa(toptax, ps.toptax_Fungi);ps.toptax_Fungi

#Class level
psmelt(ps.toptax_Fungi)
gg2=ggplot(psmelt(ps.toptax_Fungi), aes(x=reorder(Sample, Genotype), y=Abundance, fill=Class))+
  geom_bar(stat = "identity", position = "stack", width = 1) +
  #scale_fill_manual(values=c(ocean.dense(7), ocean.tempo(8)))+
  #scale_fill_manual(values=c("grey35","darkturquoise","#009E73",
   #      "Orange","#6A3D9A","#00AFBB","steelblue", "#cccda7", "#bfb77b", "7e9998", "#92B2B2", "#83bbbf", "#58a0c4","#8973b4", #"#56B4E9","#86486F","#446455", "bisque4"))+
  scale_fill_manual(values = mycolors)+
  xlab("") + ylab("Relative abundance (%)") +
  facet_wrap(~reorder(Genotype, Sample), scales = "free_x", nrow = 1) + 
  ggtitle("Class, top 500 ASVs") +
  theme_pubclean() + theme(legend.position="bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0)) +
theme(legend.text = element_text(face = c(rep("italic", 5), rep("plain", 5))))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, nrow = 2))+
  theme(text=element_text(size=8,  family="Arial"))+ labs(tag = "");gg2

library(ggpubr)
tiff("Composition_Genotype_fungi.tiff", units="in", width=10, height=5, res=300)
Composition_Genotype_fungi.tiff=gg2+plot_layout(nrow = 1);Composition_Genotype_fungi.tiff
ggsave("Composition_Genotype_fungi.tiff", plot = Composition_Genotype_fungi.tiff)
dev.off()


#Order level
psmelt(ps.toptax_Fungi)
gg2=ggplot(psmelt(ps.toptax_Fungi), aes(x=reorder(Sample, Genotype), y=Abundance, fill=Order))+
  geom_bar(stat = "identity", position = "stack", width = 1) +
  #scale_fill_manual(values=c(ocean.dense(7), ocean.tempo(8)))+
  #scale_fill_manual(values=c("#25725d", "#439b5b", "#7bc369", "#abd69a", "#d8e8b0", "#cccda7", "#bfb77b", "#7e9998", "#92B2B2", "#83bbbf", "#61acc1", "#58a0c4", "#7087c3","#8973b4", "#512972"))+
  scale_fill_manual(values = mycola)+
  xlab("") + ylab("Relative abundance (%)") +
  facet_wrap(~reorder(Genotype, Sample), scales = "free_x", nrow = 1) + 
  ggtitle("Order, top 500 ASVs") +
  theme_pubclean() + theme(legend.position="bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0)) +
theme(legend.text = element_text(face = c(rep("italic", 5), rep("plain", 5))))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, nrow = 3))+
  theme(text=element_text(size=8,  family="Arial"))+ labs(tag = "");gg2

library(ggpubr)
tiff("Order_Composition_Genotype_fungi.tiff", units="in", width=14, height=5, res=300)
Order_Composition_Genotype_fungi.tiff=gg2+plot_layout(nrow = 1);Order_Composition_Genotype_fungi.tiff
ggsave("Order_Composition_Genotype_fungi.tiff", plot = Order_Composition_Genotype_fungi.tiff)
dev.off()

```

#Average percentage composition class
```{r,echo=FALSE,warning=FALSE,message=FALSE}
 load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
#Convert to relative abundance or the computation of the micribial composition in the samples
#consider phylum
GPf_Phylum = tax_glom(ps3_Fungi, "Phylum") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Phylum = data.frame(Phylum = tax_table(GPf_Phylum)[,"Phylum"], Mean = rowMeans(otu_table(GPf_Phylum)), row.names = NULL)
df_Phylum = df_Phylum[order(-df_Phylum$Mean),];df_Phylum
head(df_Phylum,20)

#consider Class
GPf_Class = tax_glom(ps3_Fungi, "Class") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Class = data.frame(Class = tax_table(GPf_Class)[,"Class"], Mean = rowMeans(otu_table(GPf_Class)), row.names = NULL)
df_Class = df_Class[order(-df_Class$Mean),];df_Class
head(df_Class,20)

#consider Order
GPf_Order = tax_glom(ps3_Fungi, "Order") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Order = data.frame(Order = tax_table(GPf_Order)[,"Order"], Mean = rowMeans(otu_table(GPf_Order)), row.names = NULL)
df_Order = df_Order[order(-df_Order$Mean),];df_Order
head(df_Order,20)

#consider Family
GPf_Family = tax_glom(ps3_Fungi, "Family") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Family = data.frame(Family = tax_table(GPf_Family)[,"Family"], Mean = rowMeans(otu_table(GPf_Family)), row.names = NULL)
df_Family = df_Family[order(-df_Family$Mean),];df_Family
head(df_Family,20)

#consider Genus
GPf_Genus = tax_glom(ps3_Fungi, "Genus") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Genus = data.frame(Genus = tax_table(GPf_Genus)[,"Genus"], Mean = rowMeans(otu_table(GPf_Genus)), row.names = NULL)
df_Genus = df_Genus[order(-df_Genus$Mean),];df_Genus
head(df_Genus,20)

```



#Beta diversity

```{r}
phyloseq_transform_css <- function(physeq, norm = TRUE, log = TRUE, ...){
  # require(metagenomeSeq)
  MGS <- phyloseq::phyloseq_to_metagenomeSeq(physeq)
  # get the normalized count matrix
  otu_norm <- metagenomeSeq::MRcounts(MGS, norm = norm, log = log, ...)
  # exportMat(datt.norm, file = "tmp.txt")    # norm = TRUE, log = TRUE
  ## save sample statistics (sample scaling factor, quantile value, number of identified features and library size):
  # exportStats(dattM.sf, file = "tmp_stats.txt")
  # Substitue raw abundance to the css-normalized data
  physeq.tr <- physeq
  phyloseq::otu_table(physeq.tr) <- phyloseq::otu_table(otu_norm, taxa_are_rows = T)
  return(physeq.tr)
}
```


```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
ps3_Fungi_css = phyloseq_transform_css(ps3_Fungi);ps3_Fungi_css#########
#First get otu_table and transpose it:
dist.matrix_ps3_Fungi_css <- t(data.frame(otu_table(ps3_Fungi_css)))
sampledf_ps3_Fungi_css <- data.frame(sample_data(ps3_Fungi_css))
#Then use vegdist from vegan to generate a bray distance object:
phyloseq_bray_ps3_Fungi_css <- vegdist(dist.matrix_ps3_Fungi_css, method = "bray"); phyloseq_bray_ps3_Fungi_css
sampleda <- data.frame(sample_data(ps3_Fungi_css), check.names=FALSE);sampleda
sampleda <- sampleda[match(colnames(as.matrix(phyloseq_bray_ps3_Fungi_css)),rownames(sampleda)),,drop=FALSE];sampleda
sampleda$Sample_type <- factor(sampleda$Sample_type)
adores <- adonis2(phyloseq_bray_ps3_Fungi_css ~ Genotype, data=sampleda, permutation=999);adores#change from adonis to adonis2
#         Df SumOfSqs      R2      F Pr(>F)    
#Model     4   4.9094 0.62459 10.399  0.001 ***
#Residual 25   2.9508 0.37541                  
#Total    29   7.8602 1.00000                  
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#############################
ps3_Fungi_css = phyloseq_transform_css(ps3_Fungi);ps3_Fungi_css#########
#visualize
matrixdistance_ps3_Fungi_css_PCoA <- ordinate(physeq = ps3_Fungi_css, method = "PCoA", distance = "bray", trymax = 1000)
#Stress:     0.2069558 

pcoaplot1_fungi.tiff=plot_ordination(physeq = ps3_Fungi_css, ordination = matrixdistance_ps3_Fungi_css_PCoA, color = "Genotype") + 
  theme(text = element_text(size = 12)) + scale_color_manual(values =mycola)+theme(legend.title=element_blank())+  geom_point(aes(color =Genotype ), size = 5)+
  scale_shape_manual(values=c(19,17, 15,18))+ theme_gray(base_size = 14)+ labs(tag = "D")+
  theme(legend.position="");pcoaplot1_fungi.tiff
tiff("pcoaplot1_fungi.tiff", units="in", width=5, height=5, res=300)
ggsave("pcoaplot1_fungi.tiff", plot = pcoaplot1_fungi.tiff)
dev.off()
```



#Beta dispersion tests & plot
```{r}
dist=phyloseq_bray_ps3_Fungi_css
dispr <- vegan::betadisper(dist, phyloseq::sample_data(ps3_Fungi_css)$Genotype);dispr #calculate the dispersion

data_fram_cultivar <- data.frame(Distance_to_centroid=dispr$distances,Group=dispr$group);data_fram_cultivar
groups <- dispr$group

#stats
kruskal.test(Distance_to_centroid ~ Group, data = data_fram_cultivar)
#Kruskal-Wallis chi-squared = 12.903, df = 4, p-value = 0.01176

#Groups differ in both central tendency and variability (composition and spread)

```



#alpha diversity
```{r}
 load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")

Fungi_subsamp <- rarefy_even_depth(ps3_Fungi, sample.size=5500  , rngseed=5163, replace=FALSE, trimOTUs=TRUE);Fungi_subsamp#[ 1128 samples by 20 sample variables ]
Fungi_richness = estimate_richness(Fungi_subsamp , measures=c("Observed","Shannon"))
Fungi_richness_table <- as.data.frame(as.matrix(Fungi_richness))
write.csv(Fungi_richness_table,file = "Fungi_richness_table.csv")
Fungi_richness <- read.table("Fungi_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Fungi_richness
```


#Fungi diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Fungi_richness <- read.table("Fungi_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Fungi_richness
attach(Fungi_richness)
plot_Fungi_rich=ggboxplot(Fungi_richness, x = "Genotype", y = "Observed",fill  = "Genotype", ylab = "Observed ASVs", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_Fungi_rich


tiff("Fungi_rich.tiff", units="in", width=2, height=3, res=300)
Fungi_rich.tiff=plot_Fungi_rich+plot_layout(ncol =1 );Fungi_rich.tiff
ggsave("Fungi_rich.tiff", plot = Fungi_rich.tiff)
dev.off()



plot_Fungi_Shannon=ggboxplot(Fungi_richness, x = "Genotype", y = "Shannon",fill  = "Genotype", ylab = "Shannon index", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_Fungi_Shannon


tiff("Fungi_Shannon.tiff", units="in", width=2, height=3, res=300)
Fungi_Shannon.tiff=plot_Fungi_Shannon+plot_layout(ncol =1 );Fungi_Shannon.tiff
ggsave("Fungi_Shannon.tiff", plot = Fungi_Shannon.tiff)
dev.off()
```



#Fungi diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Fungi_richness <- read.table("Fungi_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Fungi_richness
attach(Fungi_richness)

#Consider microbial richness
#Normality
shapiro.test(Observed)
#W = 0.92561, p-value = 0.04239

#statistics measures for the different Fungi indices
kruskal.test(Observed ~ Genotype, data = Fungi_richness)

#data:  Observed by Genotype
#Kruskal-Wallis chi-squared = 23.146, df = 4, p-value = 0.0001184

#Post-hoc (Dunn's tests)


Fungi_Dunn_observed_genotype=dunn_test(Fungi_richness, Observed ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Fungi_Dunn_observed_genotype
write_csv(Fungi_Dunn_observed_genotype, "Fungi_Dunn_observed_genotype.csv")
Fungi_Dunn_observed_genotype <- dunnTest(Observed, Genotype, method = "bonferroni");Fungi_Dunn_observed_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Fungi_Dunn_observed_genotype = cldList(P.adj ~ Comparison, data=Fungi_Dunn_observed_genotype$res);CLD_Fungi_Dunn_observed_genotype 
```


#Fungi diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Fungi_richness <- read.table("Fungi_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Fungi_richness
attach(Fungi_richness)

#Consider microbial Shannon
#Normality
shapiro.test(Shannon)
#W = 0.91522, p-value = 0.02311

#statistics measures for the different Fungi indices
kruskal.test(Shannon ~ Genotype, data = Fungi_richness)

#data:  Shannon by Genotype
#Kruskal-Wallis chi-squared = 25.005, df = 4, p-value = 5.02e-05

#Post-hoc (Dunn's tests)


Fungi_Dunn_Shannon_genotype=dunn_test(Fungi_richness, Shannon ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Fungi_Dunn_Shannon_genotype
write_csv(Fungi_Dunn_Shannon_genotype, "Fungi_Dunn_Shannon_genotype.csv")
Fungi_Dunn_Shannon_genotype <- dunnTest(Shannon, Genotype, method = "bonferroni");Fungi_Dunn_Shannon_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Fungi_Dunn_Shannon_genotype = cldList(P.adj ~ Comparison, data=Fungi_Dunn_Shannon_genotype$res);CLD_Fungi_Dunn_Shannon_genotype 

```



#Fungi abundance

```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Fungil_Abundance <- read.table("Fungi_abundance.txt", header=TRUE, sep="\t"); Fungil_Abundance

attach(Fungil_Abundance)
variable.names(Fungil_Abundance)
Fungil_Abundance$log10
head(Fungil_Abundance)
str(Fungil_Abundance)

plot_Fungil_Abundance=ggboxplot(Fungil_Abundance, x = "Genotype", y = "log10",fill  = "Genotype", ylab = "Log(10) gene copies/gram", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_Fungil_Abundance


tiff("Fungil_abundance.tiff", units="in", width=2, height=3, res=300)
Fungil_abundance.tiff=plot_Fungil_Abundance+plot_layout(ncol =1 );Fungil_abundance.tiff
ggsave("Fungil_abundance.tiff", plot = Fungil_abundance.tiff)
dev.off()



#statistics
#Normality
shapiro.test(log10)
#W = 0.85632, p-value = 0.001029

#statistics measures for the different Fungil indices
kruskal.test(log10 ~ Genotype, data = Fungil_Abundance)

#data:  log10 by Genotype
#Kruskal-Wallis chi-squared = 26.184, df = 4, p-value = 2.905e-05

#Post-hoc (Dunn's tests)


Fungil_Dunn_log10_genotype=dunn_test(Fungil_Abundance, log10 ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Fungil_Dunn_log10_genotype
write_csv(Fungil_Dunn_log10_genotype, "Fungil_Dunn_log10_genotype.csv")
Fungil_Dunn_log10_genotype <- dunnTest(log10, Genotype, method = "bonferroni");Fungil_Dunn_log10_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Fungil_Dunn_log10_genotype = cldList(P.adj ~ Comparison, data=Fungil_Dunn_log10_genotype$res);CLD_Fungil_Dunn_log10_genotype 

```





#Differentially abundant Fungal taxa

```{r}
#https://rpubs.com/kjojess/1013527
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
#Install corncob using:
#remotes::install_github("statdivlab/corncob")
library(corncob)
library(lefser)
require(dplyr)
require(tibble)
library(qiime2R)
library(circlize)
library(ggplot2)
library(microbiomeMarker)
library(RColorBrewer)
#library(ggthemr)
library(plyr)
library(phyloseq)
library(tidyverse)
library(ComplexHeatmap)
library(cowplot)
library(phyloseq)
library(microbiomeMarker)
#Load the data
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
set.seed(1)
#sample_data(ps3_Fungi)

setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

lefse_cc_path_Genus_fungi <- run_lefse(
    ps3_Fungi,
    wilcoxon_cutoff = 0.05,
    group = "Genotype",
    kw_cutoff = 0.05,
    multigrp_strat = TRUE,
    lda_cutoff =4,
    taxa_rank="Genus"
)
lefse_cc_path_Genus_fungi
sig_marker_lefse<-data.frame(marker_table(lefse_cc_path_Genus_fungi))

plot_abundance(lefse_cc_path_Genus_fungi, group="Genotype")
Lefse_fungi=plot_ef_bar(lefse_cc_path_Genus_fungi);Lefse_fungi

```

#Assemble the differentially abundant taxa
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

library(ggpubr)
tiff("Differential_abundance.tiff", units="in", width=8, height=6, res=300)
Differential_abundance.tiff= Lefse_bacteria+Lefse_fungi+plot_layout(nrow= 2);Differential_abundance.tiff
ggsave("Differential_abundance.tiff", plot = Differential_abundance.tiff)
dev.off()

```


#Random forest classifiers of plant genpotype (using the seed microbiome)
#Load packages 
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
# Define prevalence of each taxa; # (in how many samples did each taxa appear at least once)
library(doParallel);packageVersion("doParallel")#‘1.0.17’
library(foreach);packageVersion("foreach")#‘1.5.2’
library(tidyverse);packageVersion("tidyverse")#‘2.0.0’
library(ggplot2);packageVersion("ggplot2")#‘3.5.0’
library(vegan);packageVersion("vegan")#‘2.6.4’
library(reshape2);packageVersion("reshape2")#‘1.4.4’
library(phyloseq);packageVersion("phyloseq")#‘1.46.0’
library(randomForest);packageVersion("randomForest")
library(doBy);packageVersion("doBy")#‘4.7.1.1’
library(plotrix);packageVersion("plotrix")#‘3.8.4’
library(ggpubr);packageVersion("ggpubr")#‘0.6.0’

Fungi_otu <- read.table("Fungi_Table-rf.txt", header=TRUE,row.names="OTUID");Fungi_otu
Fungi_otu <- otu_table(Fungi_otu, taxa_are_rows = TRUE);Fungi_otu 

## Imported taxonomy
#Fungi_tax <- read.csv("Taxonomy.csv",header=TRUE,row.names="OTUID");Fungi_tax
Fungi_tax <- read.table("Fungi_Taxonomy-rf.txt", header=TRUE, sep="\t",row.names="OTUID");Fungi_tax
Fungi_tax <- as.matrix(Fungi_tax);Fungi_tax
Fungi_tax <- tax_table(Fungi_tax);Fungi_tax

## Imported metadata file
#metadata <- read.table("Metadata.txt", header=TRUE, sep="\t",row.names="SampleID");metadata
Fungi_metadata<- read.csv("Fungi_metadata.csv",header=TRUE,row.names="SampleID");Fungi_metadata

#rownames(Fungi_metadata) <- Fungi_metadata[,1]
Fungi_metadata <- sample_data(Fungi_metadata);Fungi_metadata
#tree
#tree <- read.tree(file = "tree.nwk");tree
#tree <- phy_tree(tree)

all(rownames(Fungi_metadata) %in% colnames(Fungi_otu))
#Convert to phyloseq object
Fungi <- merge_phyloseq(Fungi_otu,Fungi_tax, Fungi_metadata);Fungi#[ 16460 taxa and 1197 samples ]
Fungi = subset_taxa(Fungi, Kingdom=="Fungi");Fungi# [ 6448 taxa and 84 samples ]
Fungi = subset_taxa(Fungi, Phylum!="CyanoFungi");Fungi#  [ 15588 taxa and 1197 samples ]
Fungi = subset_taxa(Fungi, Order!="Chloroplast");Fungi#  [ 15588 taxa and 1197 samples ]
Fungi = subset_taxa(Fungi, Family!="Mitochondria");Fungi# [ 15588 taxa and 1197 samples ]
Fungi <- prune_taxa(taxa_sums(Fungi)>0, Fungi);Fungi# [ 15588 taxa and 1197 samples ]
#OTU_table_Fungi = as(otu_table(Fungi), "matrix")
#########################################################################################################
#Summarize 
microbiome::summarize_phyloseq(Fungi)
#Number of singletons = 68"
#Sparsity = 0.995179763625127"
#Average number of reads = 22794.9197994987"
#Total number of reads = 27285519"
#Max. number of reads = 225892"
#Min. number of reads = 68"
sums_Fungi <- sample_sums(Fungi);sums_Fungi
sums_Fungi <- data.frame(sums_Fungi);sums_Fungi
sums_Fungi <- as.matrix(sums_Fungi);sums_Fungi
colSums(sums_Fungi) #  27281011      
max(sums_Fungi) #225892
min(sums_Fungi) # 68
sums_Fungi
#https://david-barnett.github.io/microViz/articles/web-only/tax-fixing.html #fixing the table
#https://rpubs.com/bioguo/680490 (Random forest classifications)
library(devtools);packageVersion("devtools")# ‘2.4.5’# Load the devtools package
#install_github("guokai8/microbial") # Install the package (Microbial package)
library(microbial);packageVersion("microbial")#‘0.0.22’####
ps_RandomForest_Fungi=Fungi
save(ps_RandomForest_Fungi,
file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps_RandomForest_Fungi.RData")


#Fungil claasifier
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps_RandomForest_Fungi.RData")

prunescale = 0.05

# Prune out rare OTUs by mean relative abundance set by prunescale
tax.mean.ps_RandomForest_Fungi <- taxa_sums(ps_RandomForest_Fungi)/nsamples(ps_RandomForest_Fungi) # average number of reads for each otu
sites.prune.ps_RandomForest_Fungi <- prune_taxa(tax.mean.ps_RandomForest_Fungi > prunescale, ps_RandomForest_Fungi)

#Next, we are going to format a dataframe of predictors (OTUs) and responses (States)
# Make a dataframe of training data with OTUs as column and samples as rows
predictors.ps_RandomForest_Fungi<- t(otu_table(sites.prune.ps_RandomForest_Fungi));predictors.ps_RandomForest_Fungi
dim(predictors.ps_RandomForest_Fungi)
# 1197 1655

# Make one column for our outcome/response variable 
response.ps_RandomForest_Fungi <- as.factor(sample_data(sites.prune.ps_RandomForest_Fungi)$Genotype);response.ps_RandomForest_Fungi
# Combine them into 1 data frame
rf.data.ps_RandomForest_Fungi <- data.frame(response.ps_RandomForest_Fungi, predictors.ps_RandomForest_Fungi);rf.data.ps_RandomForest_Fungi
str(rf.data.ps_RandomForest_Fungi)
# Set a random seed
set.seed(5000)
# My computer has 8 processor cores - adjust parameter of the following command as needed/desired
registerDoParallel(3)

# %dopar% implemented with foreach implements each loop call as an independent function call
# and then parallelizes all the function calls across the number of processers set above with
# registerDoParallel(). First, we need to set up a dataframe to store each bootstrap RF result.
boot.imp.s <- data.frame(predictors=NULL, Try=NULL, MeanDecreaseGini=NULL)
boot.imp.s <- foreach (Try=1:100, .combine=rbind) %dopar% {
  
  ps_RandomForest_Fungi.classify <- randomForest(response.ps_RandomForest_Fungi ~., data = rf.data.ps_RandomForest_Fungi, ntree = 500, proximity=T, mtry=200)
  print(ps_RandomForest_Fungi.classify)
  
  # Make a data frame with predictor names and their importance
  imp.s <- importance(ps_RandomForest_Fungi.classify)
  imp.s <- data.frame(predictors = rownames(imp.s), imp.s)
  
  cbind(imp.s, data.frame(Try=Try)) #)
  
}

# Same (and independent) as above but this time we are storing results from the
# out-of-bag error rates, so we car report the correct-incorrect classification
# rate for each group, and what types of misclassifications were made.
boot.oob <- NULL
boot.oob <- foreach (Try=1:100, .combine=rbind) %dopar% {
  
  ps_RandomForest_Fungi.classify <- randomForest(response.ps_RandomForest_Fungi ~., data = rf.data.ps_RandomForest_Fungi, ntree = 500, proximity=T, mtry=200)
  print(ps_RandomForest_Fungi.classify)
  
  # Make a data frame with predictor names and their importance
  imp.s <- importance(ps_RandomForest_Fungi.classify)
  imp.s <- data.frame(predictors = rownames(imp.s), imp.s)
  
  c(boot.oob, ps_RandomForest_Fungi.classify$err.rate[500,1])
}

#Display and visualize results
# Look at the results
mean(boot.oob)#0.1007352
sd(boot.oob)#0.003860804
range(boot.oob)# 0.09106099 0.11027569
# Create a dataframe with summary statistics from our bootstrap runs
summary.boot.s <- summaryBy(MeanDecreaseGini ~ predictors, data= boot.imp.s, FUN = c(mean, sd, std.error));summary.boot.s
write.csv(summary.boot.s,"summary.boot.s.csv")
# Look at it
head(summary.boot.s)
str(summary.boot.s)
# Order the predictor levels by importance
imp.s.sort <- arrange(summary.boot.s, desc(MeanDecreaseGini.mean));imp.s.sort
imp.s.sort$predictors <- factor(imp.s.sort$predictors, levels = imp.s.sort$predictors)
# Select the top 50 predictors
imp.s.5 <- imp.s.sort[1:20, ];imp.s.5

# What are those OTUs?
# otunames.s = otunames of predictors
# r.s = A logical containing which rows from phyloseq object were recovered as top 40 predictors
otunames.s <- imp.s.5$predictors;otunames.s
r.s <- rownames(tax_table(ps_RandomForest_Fungi)) %in% otunames.s;r.s

# Use r.s to pull out the taxonomic information for those OTUS in otunames.s
t.table <- as.data.frame(tax_table(ps_RandomForest_Fungi)[r.s, ])
write.csv(t.table, "RF.ps_RandomForest_Fungi.t.table.csv", quote=F,row.names=T)


#VISUALIZATION HERE
# Reformat the taxonomic names and paste them to OTU names for axis labels
# (highest determined taxonomic rate, get rid of special characters, etc.)
axislabels <- with(t.table, paste(rownames(t.table), gsub("[sgfocp]__","",Genus,perl=T) %>% gsub("_"," ",.,perl=T) %>% gsub("[\\s]?(unclassified|ge)[\\s]?","",.,perl=T)))
names(axislabels) <- rownames(t.table)
i <- grep("(uncultured)|([01234567890-]$)+", axislabels, perl=T)
axislabels[i]<-paste(names(axislabels)[i], t.table$Family[i])

# Now we sum the abundance of each OTU by state so we can order the
# RF predictor OTUs by which state they were most abundant in, followed
# by their importance value. This makes the heatmap easier to interpret.
sums<-summaryBy(. ~ response.ps_RandomForest_Fungi, data = rf.data.ps_RandomForest_Fungi[, c('response.ps_RandomForest_Fungi', as.character(otunames.s))], FUN = sum, keep.names = T);sums
state.organize <- sapply(as.character(otunames.s), FUN = function(x) which.max(sums[,x]));state.organize
sums$response.ps_RandomForest_Fungi
rf.data.ps_RandomForest_Fungi$response.ps_RandomForest_Fungi
# hm is a dataframe of the abundance of the predictors, standardized and log transformed
hm <- rf.data.ps_RandomForest_Fungi[,as.character(otunames.s)] %>% decostand(method='max', MARGIN=2) %>% decostand(method='log');hm
hm$sample <- gsub('_',' ',sample_data(ps_RandomForest_Fungi)[rownames(hm),'Genotype']$Genotype)

#hm.melt <- melt(hm[c(order(state.organize), 6)]);hm.melt
hm.melt <-  as.data.table(hm[c(order(state.organize), 21)]);hm.melt
hm.melt <- melt(hm.melt, id.vars = "sample")
names(hm.melt) <- c("Tree","Taxon","StandardizedAbundance")

# Turn predictors into a factor (re-ordered the same as the hm.melt object so they match up)
imp.s.5$predictors <- factor(imp.s.5$predictors, levels(imp.s.5$predictors)[order(state.organize)])


# Creat the heatmap with the taxon names and OTUs
plot1_Fungi <- ggplot(hm.melt, aes(x=Tree, y=Taxon, fill=StandardizedAbundance)) + 
  geom_tile() + scale_fill_gradient(low="white", high="#CC79A7") + 
  theme(axis.text.x = element_text(angle=90, size=14), legend.position = "none", axis.text.y = element_text(hjust=0))  + 
  scale_y_discrete(labels=axislabels)+theme(axis.title.x = element_text(face="bold", colour="black", size=14),
                                            axis.title.y = element_text(face="bold", colour="black", size=14),
                                            axis.text.x = element_text(face="bold", colour="black", size=14),
                                            axis.text.y  = element_text(face="bold", colour="black", size=14))+ labs(tag = "",size=14);plot1_Fungi

#Creat the bargraph
plot2_Fungi <- ggplot(imp.s.5, aes(x = predictors, y = MeanDecreaseGini.mean)) +
  geom_bar(stat = "identity", fill = "grey35") +
  ylab("Bootstrap Mean \u00b1 SD Gini Index") +
  theme_bw() +
  geom_errorbar( aes(x=predictors, ymin=MeanDecreaseGini.mean-MeanDecreaseGini.sd, ymax=MeanDecreaseGini.mean+MeanDecreaseGini.sd), width=0.75, alpha=0.9, size=0.5) +
  coord_flip() + theme(axis.title.x = element_text(face="bold", colour="black", size=14),
                       axis.title.y = element_text(face="bold", colour="black", size=14),
                       axis.text.x = element_text(face="bold", colour="black", size=14),
                       axis.text.y  = element_text(face="bold", colour="black", size=14))+ labs(tag = "",size=14);plot2_Fungi
g2_Fungi <- ggplotGrob(plot2_Fungi);g2_Fungi

library(ggpubr)
library(patchwork)

#Put them together

tiff("RF1_Fungi.tiff", units="in", width=15, height=6, res=300)
RF1_Fungi.tiff=plot1_Fungi+plot2_Fungi+plot_layout(nrow=1);RF1_Fungi.tiff
ggsave("RF1_Fungi.tiff", plot = RF1_Fungi.tiff)
dev.off()
```


#Archaeal community




#Load data (ASV table, taxonomy & tree, & metadata_Archaea)
```{r,echo = F}
##Imported feature table
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

asv_tab_Archaea <- read.table("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Archaea_Table.txt", 
              header=TRUE,
              row.names="OTUID",sep="\t");asv_tab_Archaea
asv_tab_Archaea <- otu_table(asv_tab_Archaea, taxa_are_rows = TRUE);asv_tab_Archaea 

## Imported taxonomy table
tax_tab_Archaea <- read.table("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Archaea_Taxonomy.txt", 
              header=TRUE,
              sep="\t",
              row.names="OTUID");tax_tab_Archaea
tax_tab_Archaea <- as.matrix(tax_tab_Archaea);tax_tab_Archaea
tax_tab_Archaea <- tax_table(tax_tab_Archaea);tax_tab_Archaea

## Imported metadata_Archaea file
metadata_Archaea <- read.csv("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Archaea_Metadata.csv",
              header=TRUE,
              row.names="SampleID");metadata_Archaea
#rownames(metadata_Archaea) <- metadata_Archaea[,1]
metadata_Archaea <- sample_data(metadata_Archaea);metadata_Archaea

#import reference sequences
## Import reference sequences
# rs_table <- Biostrings::readDNAStringSet(
#   "outputs/qiime2/dna-sequences.fasta",
#   format = "fasta",
#   nrec = -1L,
#   skip = 0L,
#   seek.first.rec = F,
#   use.names = T,
#   with.qualities = F)

#proof that the data are of equal dimensions
all(rownames(metadata_Archaea) %in% colnames(asv_tab_Archaea))

#Convert to phyloseq object
ps_Archaea <- merge_phyloseq(
  asv_tab_Archaea,
  tax_tab_Archaea, 
  metadata_Archaea);ps_Archaea#[ 16460 taxa & 1197 samples ]
summarize_phyloseq(ps_Archaea)
ps0_Archaea <- ps_Archaea # 16460 taxa & 1197 samples
save(ps0_Archaea,
file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps0_Archaea.RData")

```


#clean the data
```{r,echo = FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps0_Archaea.RData")

ps_clean_Archaea = subset_taxa(ps0_Archaea, Kingdom=="Archaea");ps_clean_Archaea# [ 6448 taxa & 84 samples ]
#ps_clean_Archaea = subset_taxa(ps_clean_Archaea, Phylum!="CyanoArchaea");ps_clean_Archaea#  [ 15588 taxa & 1197 samples ]
#ps_clean_Archaea = subset_taxa(ps_clean_Archaea, Order!="Chloroplast");ps_clean_Archaea#  [ 15588 taxa & 1197 samples ]
#ps_clean_Archaea = subset_taxa(ps_clean_Archaea, Family!="Mitochondria");ps_clean_Archaea# [ 15588 taxa & 1197 samples ]
ps_clean_Archaea <- prune_taxa(taxa_sums(ps_clean_Archaea)>0, ps_clean_Archaea);ps_clean_Archaea# [ 15588 taxa & 1197 samples ]

ps1_Archaea <- ps_clean_Archaea

sort(sample_sums(ps1_Archaea))
summary(sample_sums(ps1_Archaea)) # mean reads/sample = 22791
summarize_phyloseq(ps1_Archaea) # Average reads = 22791, Total reads = 27281011, Singletons = 38
taxa_names(ps1_Archaea)[1:5]
sample_data(ps1_Archaea)$total_reads <- sample_sums(ps1_Archaea)

save(ps1_Archaea, file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps1_Archaea.RData")

K <- as.data.frame(table(tax_table(ps1_Archaea)[,"Kingdom"]));K#56
P <- as.data.frame(table(tax_table(ps1_Archaea)[,"Phylum"]));P # 4 phyla
C <- as.data.frame(table(tax_table(ps1_Archaea)[,"Class"]));C # 5 classes
O <- as.data.frame(table(tax_table(ps1_Archaea)[,"Order"]));O # 5 orders
Fa <- as.data.frame(table(tax_table(ps1_Archaea)[,"Family"]));Fa # 7 families
G <- as.data.frame(table(tax_table(ps1_Archaea)[,"Genus"]));G # 20 genera
#S <- as.data.frame(table(tax_table(ps1_Archaea)[,"Species"]));S # 1264 species

microbiome::summarize_phyloseq(ps1_Archaea)
#Number of singletons = 68"
#Sparsity = 0.995179763625127"
#Average number of reads = 22794.9197994987"
# Total number of reads = 27281011"
#Max. number of reads = 225892"
#Min. number of reads = 68"
taxonomy <- as.vector(phyloseq::tax_table(ps1_Archaea))
sums_ps1_Archaea <- sample_sums(ps1_Archaea);sums_ps1_Archaea
sums_ps1_Archaea <- data.frame(sums_ps1_Archaea);sums_ps1_Archaea
sums_ps1_Archaea <- as.matrix(sums_ps1_Archaea);sums_ps1_Archaea
colSums(sums_ps1_Archaea) #  27281011      
max(sums_ps1_Archaea) #225892
min(sums_ps1_Archaea) # 68
sums_ps1_Archaea
```


#Filter data at 0.025 prevalence

#Obtained from: https://f1000research.com/articles/5-1492/v1;#One of the reasons to filter by prevalence is to avoid spending #much time analyzing taxa that were only rarely seen. This also turns out to be a useful filter of noise (taxa that are #actually just artifacts of the data collection process)
# Define prevalence of each taxa; # (in how many samples did each taxa appear at least once)

```{r,echo = FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps1_Archaea.RData")

prev0 = apply(X = otu_table(ps1_Archaea),
              MARGIN = ifelse(taxa_are_rows(ps1_Archaea), yes = 1, no = 2),
              FUN = function(x){sum(x > 0)})

prevdf = data.frame(Prevalence = prev0,
                    TotalAbundance = taxa_sums(ps1_Archaea),
                    tax_table(ps1_Archaea))# the most prevalent ASVs present therein the composition

#only for reference, in case we want to filter Phyla that appear less than X times
keepPhyla = table(prevdf$Phylum)[(table(prevdf$Phylum) > 0)]

prevdf1 = subset(prevdf, Phylum %in% names(keepPhyla))

# Define prevalence threshold as 0.25% of total samples# 
prevalenceThreshold = 0.00025 * nsamples(ps1_Archaea)
prevalenceThreshold

# Execute prevalence filter, using `prune_taxa()` function
ps2_Archaea_Archaea = prune_taxa((prev0 > prevalenceThreshold), ps1_Archaea);ps2_Archaea_Archaea

# Filter entries with unidentified Phylum.
ps3_Archaea = subset_taxa(ps2_Archaea_Archaea, Phylum %in% names(keepPhyla));ps3_Archaea

#Plot prevalence of different phyla
#https://github.com/joey711/phyloseq/issues/1474

ggplot(prevdf1, aes(TotalAbundance, Prevalence, color = Phylum)) +
  geom_hline(yintercept = prevalenceThreshold, alpha = 0.5, linetype = 2) +
  geom_point(size = 2, alpha = 0.65) +
  scale_y_log10() + scale_x_log10() +
  ggtitle(paste0("Prevalence filtering, using a threshold of ", prevalenceThreshold))+
  xlab("Total Abundance") +
  scale_color_manual(values=ocean.balance(38))+
  theme_bw()+
  facet_wrap(~Phylum, nrow = 4)

print(paste0("Initial #ASVs:", ntaxa(ps1_Archaea), ". #ASVs after prevalence filtering:", ntaxa(ps3_Archaea), ". #ASVs removed: ", ntaxa(ps1_Archaea)-ntaxa(ps3_Archaea)))
```




## Ab-Prev prep
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")

meta <- microbiome::meta(ps3_Archaea) # x_bac
asv <- otu_table(ps3_Archaea) # x_bac
taxo <- rownames_to_column(as.data.frame(tax_table(ps3_Archaea)),
                                        "ASV") # x_bac
sample_names(ps3_Archaea)
# presence-absence data
asv_PA <- 1 * ((asv > 0) == 1)

# occupancy calculation
asv_prev <- rowSums(asv_PA) / ncol(asv_PA)

# relative abundance
asv_rel <- apply(vegan::decostand(asv, 
                                  method = "total", 
                                  MARGIN = 2), 1, mean)

# combine occupancy and RA of each ASV in a table
asv_prev_rel <- rownames_to_column(as.data.frame(cbind(asv_prev, 
                                                       asv_rel)),
                                   "ASV")

asv_prev_rel_taxo <- merge(taxo, 
                           asv_prev_rel, 
                           by = "ASV")

# merge with metadata
asv_tdf <- rownames_to_column(as.data.frame(t(asv)),
                           "replicates")
library(dplyr)
meta <- meta <- meta %>% rename_with(~ "replicates", 1)
colnames(meta)

colnames(asv_tdf)
asv_tdf_meta <- merge(meta,
                      asv_tdf, 
                      by = "replicates")

# keep only the one feature column for collaps3_Archaeaing table
asv_tdf_meta <- asv_tdf_meta[-c(1, 3:29)]

asv_gen <- asv_tdf_meta %>% 
    group_by(Genotype) %>%
  summarise_each(list(sum))

# make a ASV P/A table
asv_gen_PA <- 1 * ((asv_gen > 0) == 1)

# merge prevalence, RA and taxa with the nr of obs per genotype
asv_gen_obs <- data.frame(colSums(asv_gen_PA)) %>%
  na.omit() %>% 
  rownames_to_column("ASV")
names(asv_gen_obs)[names(asv_gen_obs) == "colSums.asv_gen_PA."] <- 'Obs'

asv_prev_rel_taxo_gen_obs <- merge(asv_prev_rel_taxo, 
                                   asv_gen_obs, 
                                   by = "ASV")
```


## Ab-Prev Plot 
```{r}
counts <- asv_prev_rel_taxo_gen_obs %>% dplyr::count(Kingdom = "Archaea")

p_core_flex1_all_Archaea <- 
  ggplot(asv_prev_rel_taxo_gen_obs, 
         aes(x = asv_rel, y = asv_prev, size = Obs)) +
  geom_point(
    color = "darkgrey",
    alpha = 0.5) + 
  ggrepel::geom_text_repel(data = subset(asv_prev_rel_taxo_gen_obs, asv_prev > 0.5),
                           max.overlaps3_Archaea = 20,
                           aes(asv_rel,
                               asv_prev,
                               label = Genus),
                           size = 3,
                           show.legend = F) +
  geom_text(data = counts,
            aes(x = 0.00000001, y = 0.1, 
                label = paste0("Total ASVs\n", n),  
                fontface = "bold"), size = 4) +
  geom_hline(yintercept = 0.9, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.7, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.5, linetype = "dotted", color = "red") +
  geom_hline(yintercept = 0.25, linetype = "dotted", color = "red") +
  geom_vline(xintercept = 0.00005, linetype = "dotted", color = "red", show.legend = T) +
  theme_classic() +
  theme(
    plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm"),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.line = element_blank(),
    legend.text = element_text(size = 12),
    legend.position = c(0.35, 0.95),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "transparent", color = "transparent"),
    legend.margin = margin(0, 0, 0, 0),
    panel.spacing = unit(0.1, "lines")) +
  scale_x_log10(breaks = c(0.1, 0.01, 0.001, 0.0001, 0.00001, 0.000001, 0.0000001), 
                labels = c(10, 1, 0.1, 0.01, 0.001, 0.0001, 0.00001)) +
  scale_y_continuous(limits = c(-0.01, 1.01),
                     breaks = c(0,0.25, 0.5, 0.7, 0.9),
                     labels = c(0,25, 50, 70, 90)) +
  labs(x = "Relative abundance [%]", y = "Prevalence [%]", size = "nº cultivars"); p_core_flex1_all_Archaea

 ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ab-prev-all_Archaea.tiff",     
  plot = p_core_flex1_all_Archaea,
   scale = 1,
 width = 25,
  height = 20,
  units = c("cm"),
 dpi = 600)
```



# Save the Archaeal object
```{r}
save(ps3_Archaea, file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
```



#comsider order level
#https://www.geeksforgeeks.org/networkd3-package-in-r/
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
library("MiscMetabar")
library("phyloseq")
library("networkD3")
#summary_plot_pq(ps3_Archaea)
#sankey_pq(gp, taxa = c(1:5))
#upset_pq(ps3_Archaea, "Genotype", taxa = "Genus")


toptax = names(sort(taxa_sums(ps3_Archaea), decreasing=TRUE))[1:100];toptax
ps.toptax_Archaea = transform_sample_counts(ps3_Archaea, function(x) {x/sum(x)*100});ps.toptax_Archaea
ps.toptax_Archaea <- prune_taxa(toptax, ps.toptax_Archaea);ps.toptax_Archaea
#dim(otu_table(ps.toptax_Archaea))
#rank_names(ps.toptax_Archaea)
#sample_variables(ps.toptax_Archaea)
ps.toptax_Archaea <- prune_samples(sample_sums(ps.toptax_Archaea) > 0, ps.toptax_Archaea);ps.toptax_Archaea
ps.toptax_Archaea <- prune_taxa(taxa_sums(ps.toptax_Archaea) > 0, ps.toptax_Archaea);ps.toptax_Archaea


if (nsamples(ps.toptax_Archaea) > 0 && ntaxa(ps.toptax_Archaea) > 0) {
  upset_Archaea_Order <- upset_pq(ps.toptax_Archaea, "Genotype", taxa = "Order")
} else {
  message("No data available for Archaea at Order level.")
}


upset_Archaea_Order=upset_pq(ps.toptax_Archaea, "Genotype", taxa = "Order");upset_Archaea_Order
ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Shared_upset_Archaea_Order.tiff",     
  plot = upset_Archaea_Order,
  scale = 1,
  width = 20,
 height = 11,
 units = c("cm"),
  dpi = 600)
```



#consider the genus level
#https://www.geeksforgeeks.org/networkd3-package-in-r/
#https://adrientaudiere.github.io/MiscMetabar/articles/MiscMetabar.html
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
library("MiscMetabar")
library("phyloseq")
library("networkD3")
#summary_plot_pq(ps3_Archaea)
#sankey_pq(gp, taxa = c(1:5))
#upset_pq(ps3_Archaea, "Genotype", taxa = "Genus")

toptax = names(sort(taxa_sums(ps3_Archaea), decreasing=TRUE))[1:100];toptax
ps.toptax_Archaea = transform_sample_counts(ps3_Archaea, function(x) {x/sum(x)*100});ps.toptax_Archaea
ps.toptax_Archaea <- prune_taxa(toptax, ps.toptax_Archaea);ps.toptax_Archaea



toptax = names(sort(taxa_sums(ps3_Archaea), decreasing=TRUE))[1:100];toptax
ps.toptax_Archaea = transform_sample_counts(ps3_Archaea, function(x) {x/sum(x)*100});ps.toptax_Archaea
ps.toptax_Archaea <- prune_taxa(toptax, ps.toptax_Archaea);ps.toptax_Archaea
#dim(otu_table(ps.toptax_Archaea))
#rank_names(ps.toptax_Archaea)
#sample_variables(ps.toptax_Archaea)
ps.toptax_Archaea <- prune_samples(sample_sums(ps.toptax_Archaea) > 0, ps.toptax_Archaea);ps.toptax_Archaea
ps.toptax_Archaea <- prune_taxa(taxa_sums(ps.toptax_Archaea) > 0, ps.toptax_Archaea);ps.toptax_Archaea


if (nsamples(ps.toptax_Archaea) > 0 && ntaxa(ps.toptax_Archaea) > 0) {
  upset_Archaea_Order <- upset_pq(ps.toptax_Archaea, "Genotype", taxa = "Genus")
} else {
  message("No data available for Archaea at Order level.")
}


upset_Archaea_Genus=upset_pq(ps.toptax_Archaea, "Genotype", taxa = "Genus");upset_Archaea_Genus

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Shared_upset_Archaea_Genus.tiff",     
 plot = upset_Archaea_Genus,
  scale = 1,
  width = 18,
 height = 11,
 units = c("cm"),
  dpi = 600)
```




##plot rarefaction curves  for the Archaeal community
#code based on https://microsud.github.io/microbiomeutilities/articles/microbiomeutilities.html
```{r, echo=FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
library(microbiomeutilities)
# set seed
set.seed(1)
subsamples <- seq(0, 50000, by=500)[-1]
#subsamples = c(10, 5000, 10000, 20000, 30000)

Rareaction_Archaea <- plot_alpha_rcurve(ps3_Archaea, index="observed",
            subsamples = subsamples,
            lower.conf = 0.025, 
            upper.conf = 0.975,
            group="Genotype",
            label.color = "brown3",
            label.size = 3,
            label.min = TRUE);Rareaction_Archaea

Rareaction_Archaea <- Rareaction_Archaea + scale_color_manual(values = mycola) + 
  scale_fill_manual(values = mycola)+ 
  theme(axis.text=element_text(size=8), panel.grid=element_blank(),
        strip.background = element_rect(colour=NA,fill="grey33"),
        strip.text.x = element_text(face="bold"))+
  theme(text=element_text(size=14,  family="sans"))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, ncol=6))+
  theme(legend.position="none")+
  labs(tag = "")+   expand_limits(x = 0, y = 0);Rareaction_Archaea

tiff("p_Archaea.tiff", units="in", width=12, height=5, res=300)
ggsave("p_Archaea.tiff", plot = Rareaction_Archaea)

```


#Transformed to relative abundances 
#bar blot of the seed microbiome composition in the different cukltivars at the various locations for the top500 ASVs

```{r,echo=FALSE,warning = FALSE}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
ps3_Archaea
# Bar plots:
##Showing individual replicates. Top 500 taxa
toptax = names(sort(taxa_sums(ps3_Archaea), decreasing=TRUE))[1:100];toptax
ps.toptax_Archaea = transform_sample_counts(ps3_Archaea, function(x) {x/sum(x)*100});ps.toptax_Archaea
ps.toptax_Archaea <- prune_taxa(toptax, ps.toptax_Archaea);ps.toptax_Archaea

#Class level
psmelt(ps.toptax_Archaea)
gg3=ggplot(psmelt(ps.toptax_Archaea), aes(x=reorder(Sample, Genotype), y=Abundance, fill=Class))+
  geom_bar(stat = "identity", position = "stack", width = 1) +
  #scale_fill_manual(values=c(ocean.dense(7), ocean.tempo(8)))+
  #scale_fill_manual(values=c("grey35","darkturquoise","#009E73",
   #      "Orange","#6A3D9A","#00AFBB","steelblue", "#cccda7", "#bfb77b", "7e9998", "#92B2B2", "#83bbbf", "#58a0c4","#8973b4", #"#56B4E9","#86486F","#446455", "bisque4"))+
  scale_fill_manual(values = mycolors)+
  xlab("") + ylab("Relative abundance (%)") +
  facet_wrap(~reorder(Genotype, Sample), scales = "free_x", nrow = 1) + 
  ggtitle("Class, top 500 ASVs") +
  theme_pubclean() + theme(legend.position="bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0)) +
theme(legend.text = element_text(face = c(rep("italic", 5), rep("plain", 5))))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, nrow = 1))+
  theme(text=element_text(size=8,  family="Arial"))+ labs(tag = "");gg3

library(ggpubr)
tiff("Composition_Genotype_Archaea.tiff", units="in", width=9, height=5, res=300)
Composition_Genotype_Archaea.tiff=gg3+plot_layout(nrow = 1);Composition_Genotype_Archaea.tiff
ggsave("Composition_Genotype_Archaea.tiff", plot = Composition_Genotype_Archaea.tiff)
dev.off()


#Order level
psmelt(ps.toptax_Archaea)
gg3=ggplot(psmelt(ps.toptax_Archaea), aes(x=reorder(Sample, Genotype), y=Abundance, fill=Order))+
  geom_bar(stat = "identity", position = "stack", width = 1) +
  #scale_fill_manual(values=c(ocean.dense(7), ocean.tempo(8)))+
  #scale_fill_manual(values=c("#25725d", "#439b5b", "#7bc369", "#abd69a", "#d8e8b0", "#cccda7", "#bfb77b", "#7e9998", "#92B2B2", "#83bbbf", "#61acc1", "#58a0c4", "#7087c3","#8973b4", "#512972"))+
  scale_fill_manual(values = mycola)+
  xlab("") + ylab("Relative abundance (%)") +
  facet_wrap(~reorder(Genotype, Sample), scales = "free_x", nrow = 1) + 
  ggtitle("Order, top 500 ASVs") +
  theme_pubclean() + theme(legend.position="bottom", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0)) +
theme(legend.text = element_text(face = c(rep("italic", 5), rep("plain", 5))))+
  guides(fill= guide_legend(keywidth = 0.6, keyheight = 0.6, nrow = 2))+
  theme(text=element_text(size=8,  family="Arial"))+ labs(tag = "");gg3

library(ggpubr)
tiff("Order_Composition_Genotype_Archaea.tiff", units="in", width=14, height=5, res=300)
Order_Composition_Genotype_Archaea.tiff=gg3+plot_layout(nrow = 1);Order_Composition_Genotype_Archaea.tiff
ggsave("Order_Composition_Genotype_Archaea.tiff", plot = Order_Composition_Genotype_Archaea.tiff)
dev.off()

```

#Average percentage composition class
```{r,echo=FALSE,warning=FALSE,message=FALSE}
 load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
#Convert to relative abundance or the computation of the micribial composition in the samples
#consider phylum
GPf_Phylum = tax_glom(ps3_Archaea, "Phylum") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Phylum = data.frame(Phylum = tax_table(GPf_Phylum)[,"Phylum"], Mean = rowMeans(otu_table(GPf_Phylum)), row.names = NULL)
df_Phylum = df_Phylum[order(-df_Phylum$Mean),];df_Phylum
head(df_Phylum,20)

#consider Class
GPf_Class = tax_glom(ps3_Archaea, "Class") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Class = data.frame(Class = tax_table(GPf_Class)[,"Class"], Mean = rowMeans(otu_table(GPf_Class)), row.names = NULL)
df_Class = df_Class[order(-df_Class$Mean),];df_Class
head(df_Class,20)

#consider Order
GPf_Order = tax_glom(ps3_Archaea, "Order") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Order = data.frame(Order = tax_table(GPf_Order)[,"Order"], Mean = rowMeans(otu_table(GPf_Order)), row.names = NULL)
df_Order = df_Order[order(-df_Order$Mean),];df_Order
head(df_Order,20)

#consider Family
GPf_Family = tax_glom(ps3_Archaea, "Family") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Family = data.frame(Family = tax_table(GPf_Family)[,"Family"], Mean = rowMeans(otu_table(GPf_Family)), row.names = NULL)
df_Family = df_Family[order(-df_Family$Mean),];df_Family
head(df_Family,20)

#consider Genus
GPf_Genus = tax_glom(ps3_Archaea, "Genus") %>% transform_sample_counts(function(x) {x * 100/sum(x)})
df_Genus = data.frame(Genus = tax_table(GPf_Genus)[,"Genus"], Mean = rowMeans(otu_table(GPf_Genus)), row.names = NULL)
df_Genus = df_Genus[order(-df_Genus$Mean),];df_Genus
head(df_Genus,20)

```


#Beta diversity

```{r}
phyloseq_transform_css <- function(physeq, norm = TRUE, log = TRUE, ...){
  # require(metagenomeSeq)
  MGS <- phyloseq::phyloseq_to_metagenomeSeq(physeq)
  # get the normalized count matrix
  otu_norm <- metagenomeSeq::MRcounts(MGS, norm = norm, log = log, ...)
  # exportMat(datt.norm, file = "tmp.txt")    # norm = TRUE, log = TRUE
  ## save sample statistics (sample scaling factor, quantile value, number of identified features and library size):
  # exportStats(dattM.sf, file = "tmp_stats.txt")
  # Substitue raw abundance to the css-normalized data
  physeq.tr <- physeq
  phyloseq::otu_table(physeq.tr) <- phyloseq::otu_table(otu_norm, taxa_are_rows = T)
  return(physeq.tr)
}
```


```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
sample_sums(ps3_Archaea)
ps3_Archaea <- prune_samples(sample_sums(ps3_Archaea) > 0, ps3_Archaea);ps3_Archaea
ps3_Archaea_css = phyloseq_transform_css(ps3_Archaea);ps3_Archaea_css#########
#First get otu_table and transpose it:
dist.matrix_ps3_Archaea_css <- t(data.frame(otu_table(ps3_Archaea_css)))
sampledf_ps3_Archaea_css <- data.frame(sample_data(ps3_Archaea_css))
#Then use vegdist from vegan to generate a bray distance object:
phyloseq_bray_ps3_Archaea_css <- vegdist(dist.matrix_ps3_Archaea_css, method = "bray"); phyloseq_bray_ps3_Archaea_css
sampleda <- data.frame(sample_data(ps3_Archaea_css), check.names=FALSE);sampleda
sampleda <- sampleda[match(colnames(as.matrix(phyloseq_bray_ps3_Archaea_css)),rownames(sampleda)),,drop=FALSE];sampleda
adores <- adonis2(phyloseq_bray_ps3_Archaea_css ~ Genotype, data=sampleda, permutation=999);adores#change from adonis to adonis2
#Model     4   2.2618 0.23438 1.7603  0.012 *
#Residual 23   7.3882 0.76562                
#Total    27   9.6500 1.00000                
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
#############################
ps3_Archaea_css = phyloseq_transform_css(ps3_Archaea);ps3_Archaea_css#########
#visualize
matrixdistance_ps3_Archaea_css_PCoA <- ordinate(physeq = ps3_Archaea_css, method = "PCoA", distance = "bray", trymax = 1000)
#Stress:     0.2069558 

PCOA_Archaea_cultitvars.tiff=plot_ordination(physeq = ps3_Archaea_css, ordination = matrixdistance_ps3_Archaea_css_PCoA, color = "Genotype") + 
  theme(text = element_text(size = 12)) + scale_color_manual(values =mycola)+theme(legend.title=element_blank())+  geom_point(aes(color =Genotype ), size = 5)+
  scale_shape_manual(values=c(19,17, 15,18))+ theme_gray(base_size = 14)+ labs(tag = "E")+
  theme(legend.position="");PCOA_Archaea_cultitvars.tiff
tiff("PCOA_Archaea_cultitvars.tiff", units="in", width=6, height=3, res=300)
ggsave("PCOA_Archaea_cultitvars.tiff", plot = PCOA_Archaea_cultitvars.tiff)
dev.off()


#Combine the figures
library(patchwork)
#tiff("Beta_bact_fung_arch.tiff", units="in", width=10, height=5, res=300)
#Beta_bact_fung_arch.tiff=pcoaplot1_bacteria.tiff+pcoaplot1_fungi.tiff+PCOA_Archaea_cultitvars.tiff+plot_layout(nrow=1);Beta_bact_fung_arch.tiff
#ggsave("Beta_bact_fung_arch.tiff", plot = Beta_bact_fung_arch.tiff)
#dev.off()


```



#Beta dispersion tests & plot
```{r}
dist=phyloseq_bray_ps3_Archaea_css
dispr <- vegan::betadisper(dist, phyloseq::sample_data(ps3_Archaea_css)$Genotype);dispr #calculate the dispersion

data_fram_cultivar <- data.frame(Distance_to_centroid=dispr$distances,Group=dispr$group);data_fram_cultivar
groups <- dispr$group

#stats
kruskal.test(Distance_to_centroid ~ Group, data = data_fram_cultivar)
#Kruskal-Wallis chi-squared = 7.508, df = 4, p-value = 0.1114

#Groups differ in both central tendency and variability (composition and spread)

```


#alpha diversity
```{r}
 load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")

Archaea_subsamp <- rarefy_even_depth(ps3_Archaea, sample.size=45  , rngseed=5163, replace=FALSE, trimOTUs=TRUE);Archaea_subsamp#[ 1128 samples by 20 sample variables ]
Archaea_richness = estimate_richness(Archaea_subsamp , measures=c("Observed","Shannon"))
Archaea_richness_table <- as.data.frame(as.matrix(Archaea_richness))
write.csv(Archaea_richness_table,file = "Archaea_richness_table.csv")
Archaea_richness <- read.table("Archaea_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Archaea_richness
```


#Archaeal diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Archaea_richness <- read.table("Archaea_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Archaea_richness
attach(Archaea_richness)
plot_Archaea_rich=ggboxplot(Archaea_richness, x = "Genotype", y = "Observed",fill  = "Genotype", ylab = "Observed ASVs", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_Archaea_rich


tiff("Archaea_rich.tiff", units="in", width=2, height=3, res=300)
Archaea_rich.tiff=plot_Archaea_rich+plot_layout(ncol =1 );Archaea_rich.tiff
ggsave("Archaea_rich.tiff", plot = Archaea_rich.tiff)
dev.off()



plot_Archaea_Shannon=ggboxplot(Archaea_richness, x = "Genotype", y = "Shannon",fill  = "Genotype", ylab = "Shannon index", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_Archaea_Shannon


tiff("Archaea_Shannon.tiff", units="in", width=2, height=3, res=300)
Archaea_Shannon.tiff=plot_Archaea_Shannon+plot_layout(ncol =1 );Archaea_Shannon.tiff
ggsave("Archaea_Shannon.tiff", plot = Archaea_Shannon.tiff)
dev.off()
```



#Archaeal diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Archaea_richness <- read.table("Archaea_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Archaea_richness
attach(Archaea_richness)

#Consider microbial richness
#Normality
shapiro.test(Observed)
#W = 0.8717, p-value = 0.01898

#statistics measures for the different Archaeal indices
kruskal.test(Observed ~ Genotype, data = Archaea_richness)

#data:  Observed by Genotype
#Kruskal-Wallis chi-squared = 4.6505, df = 4, p-value = 0.3251

#Post-hoc (Dunn's tests)


Archaeal_Dunn_observed_genotype=dunn_test(Archaea_richness, Observed ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Archaeal_Dunn_observed_genotype
write_csv(Archaeal_Dunn_observed_genotype, "Archaeal_Dunn_observed_genotype.csv")
Archaeal_Dunn_observed_genotype <- dunnTest(Observed, Genotype, method = "bonferroni");Archaeal_Dunn_observed_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Archaeal_Dunn_observed_genotype = cldList(P.adj ~ Comparison, data=Archaeal_Dunn_observed_genotype$res);CLD_Archaeal_Dunn_observed_genotype 
```

#Archaeal diversity
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Archaea_richness <- read.table("Archaea_richness_table.txt", header=TRUE, sep = "\t",row.names="Sample"); Archaea_richness
attach(Archaea_richness)

#Consider microbial Shannon
#Normality
shapiro.test(Shannon)
#W = 0.95428, p-value = 0.4961

#statistics measures for the different Archaeal indices
kruskal.test(Shannon ~ Genotype, data = Archaea_richness)

#data:  Shannon by Genotype
#Kruskal-Wallis chi-squared = 2.338, df = 4, p-value = 0.6739

#Post-hoc (Dunn's tests)


Archaeal_Dunn_Shannon_genotype=dunn_test(Archaea_richness, Shannon ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Archaeal_Dunn_Shannon_genotype
write_csv(Archaeal_Dunn_Shannon_genotype, "Archaeal_Dunn_Shannon_genotype.csv")
Archaeal_Dunn_Shannon_genotype <- dunnTest(Shannon, Genotype, method = "bonferroni");Archaeal_Dunn_Shannon_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Archaeal_Dunn_Shannon_genotype = cldList(P.adj ~ Comparison, data=Archaeal_Dunn_Shannon_genotype$res);CLD_Archaeal_Dunn_Shannon_genotype 

```


#Archaea abundance
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

Archaea_Abundance <- read.table("Archaea_abundance.txt", header=TRUE, sep="\t"); Archaea_Abundance

attach(Archaea_Abundance)
variable.names(Archaea_Abundance)
Archaea_Abundance$log10
head(Archaea_Abundance)
str(Archaea_Abundance)

plot_Archaea_Abundance=ggboxplot(Archaea_Abundance, x = "Genotype", y = "log10",fill  = "Genotype", ylab = "Log(10) gene copies/gram", xlab = "Plant genotype")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ theme(legend.position="none") +geom_jitter(width = 0.3,alpha=0.2)+
  geom_jitter(width = 0.3,alpha=0.2)+theme(text = element_text(size = 8.5)) +scale_color_gradientn(colours = c('grey', 'pink'))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "gray33", size=0.4))+
  theme(axis.title.x = element_text(face="bold", colour="gray33", size=12),
        axis.title.y = element_text(face="bold", colour="gray33", size=12),
        axis.text.x = element_blank(),
        axis.text.y  = element_text( colour="gray33", size=12))+
  theme(text=element_text(size=12,  family="sans"))+ labs(tag = "");plot_Archaea_Abundance


tiff("Archaea_abundance.tiff", units="in", width=2, height=3, res=300)
Archaea_abundance.tiff=plot_Archaea_Abundance+plot_layout(ncol =1 );Archaea_abundance.tiff
ggsave("Archaea_abundance.tiff", plot = Archaea_abundance.tiff)
dev.off()


#statistics
#Normality
shapiro.test(log10)
#W = 0.85632, p-value = 0.001029

#statistics measures for the different Archaea indices
kruskal.test(log10 ~ Genotype, data = Archaea_Abundance)

#data:  log10 by Genotype
#Kruskal-Wallis chi-squared = 26.184, df = 4, p-value = 2.905e-05

#Post-hoc (Dunn's tests)


Archaea_Dunn_log10_genotype=dunn_test(Archaea_Abundance, log10 ~ Genotype, p.adjust.method = "bonferroni", detailed = FALSE);Archaea_Dunn_log10_genotype
write_csv(Archaea_Dunn_log10_genotype, "Archaea_Dunn_log10_genotype.csv")
Archaea_Dunn_log10_genotype <- dunnTest(log10, Genotype, method = "bonferroni");Archaea_Dunn_log10_genotype
# Obtaining significance letters
library(rcompanion)
CLD_Archaea_Dunn_log10_genotype = cldList(P.adj ~ Comparison, data=Archaea_Dunn_log10_genotype$res);CLD_Archaea_Dunn_log10_genotype 
```





#Differentially abundant archaeal taxa

```{r}
#https://rpubs.com/kjojess/1013527
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
rm(list = ls())
#Install corncob using:
#remotes::install_github("statdivlab/corncob")
library(corncob)
library(lefser)
require(dplyr)
require(tibble)
library(qiime2R)
library(circlize)
library(ggplot2)
library(microbiomeMarker)
library(RColorBrewer)
#library(ggthemr)
library(plyr)
library(phyloseq)
library(tidyverse)
library(ComplexHeatmap)
library(cowplot)
library(phyloseq)
library(microbiomeMarker)
#Load the data
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
set.seed(1)
#sample_data(ps3_Archaea)

setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")

lefse_cc_path_Genus_archaea <- run_lefse(
    ps3_Archaea,
    wilcoxon_cutoff = 0.05,
    group = "Genotype",
    kw_cutoff = 0.05,
    multigrp_strat = TRUE,
    lda_cutoff =2,
    taxa_rank="Genus"
)
lefse_cc_path_Genus_archaea
sig_marker_lefse<-data.frame(marker_table(lefse_cc_path_Genus_archaea))

plot_abundance(lefse_cc_path_Genus_archaea, group="Genotype")
Lefse_archaea=plot_ef_bar(lefse_cc_path_Genus_archaea);Lefse_archaea

```


#Random forest classifiers of plant genpotype (using the seed microbiome)
#Load packages 
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
# Define prevalence of each taxa; # (in how many samples did each taxa appear at least once)
library(doParallel);packageVersion("doParallel")#‘1.0.17’
library(foreach);packageVersion("foreach")#‘1.5.2’
library(tidyverse);packageVersion("tidyverse")#‘2.0.0’
library(ggplot2);packageVersion("ggplot2")#‘3.5.0’
library(vegan);packageVersion("vegan")#‘2.6.4’
library(reshape2);packageVersion("reshape2")#‘1.4.4’
library(phyloseq);packageVersion("phyloseq")#‘1.46.0’
library(randomForest);packageVersion("randomForest")
library(doBy);packageVersion("doBy")#‘4.7.1.1’
library(plotrix);packageVersion("plotrix")#‘3.8.4’
library(ggpubr);packageVersion("ggpubr")#‘0.6.0’

Archaea_otu <- read.table("Archaea_Table-rf.txt", header=TRUE,row.names="OTUID");Archaea_otu
Archaea_otu <- otu_table(Archaea_otu, taxa_are_rows = TRUE);Archaea_otu 

## Imported taxonomy
#Archaea_tax <- read.csv("Taxonomy.csv",header=TRUE,row.names="OTUID");Archaea_tax
Archaea_tax <- read.table("Archaea_Taxonomy-rf.txt", header=TRUE, sep="\t",row.names="OTUID");Archaea_tax
Archaea_tax <- as.matrix(Archaea_tax);Archaea_tax
Archaea_tax <- tax_table(Archaea_tax);Archaea_tax

## Imported metadata file
#metadata <- read.table("Metadata.txt", header=TRUE, sep="\t",row.names="SampleID");metadata
Archaea_metadata<- read.csv("Archaea_metadata.csv",header=TRUE,row.names="SampleID");Archaea_metadata

#rownames(Archaea_metadata) <- Archaea_metadata[,1]
Archaea_metadata <- sample_data(Archaea_metadata);Archaea_metadata
#tree
#tree <- read.tree(file = "tree.nwk");tree
#tree <- phy_tree(tree)

all(rownames(Archaea_metadata) %in% colnames(Archaea_otu))
#Convert to phyloseq object
Archaea <- merge_phyloseq(Archaea_otu,Archaea_tax, Archaea_metadata);Archaea#[ 16460 taxa and 1197 samples ]
Archaea = subset_taxa(Archaea, Kingdom=="Archaea");Archaea# [ 6448 taxa and 84 samples ]
Archaea = subset_taxa(Archaea, Phylum!="CyanoArchaea");Archaea#  [ 15588 taxa and 1197 samples ]
Archaea = subset_taxa(Archaea, Order!="Chloroplast");Archaea#  [ 15588 taxa and 1197 samples ]
Archaea = subset_taxa(Archaea, Family!="Mitochondria");Archaea# [ 15588 taxa and 1197 samples ]
Archaea <- prune_taxa(taxa_sums(Archaea)>0, Archaea);Archaea# [ 15588 taxa and 1197 samples ]
#OTU_table_Archaea = as(otu_table(Archaea), "matrix")
#########################################################################################################
#Summarize 
microbiome::summarize_phyloseq(Archaea)
#Number of singletons = 68"
#Sparsity = 0.995179763625127"
#Average number of reads = 22794.9197994987"
#Total number of reads = 27285519"
#Max. number of reads = 225892"
#Min. number of reads = 68"
sums_Archaea <- sample_sums(Archaea);sums_Archaea
sums_Archaea <- data.frame(sums_Archaea);sums_Archaea
sums_Archaea <- as.matrix(sums_Archaea);sums_Archaea
colSums(sums_Archaea) #  27281011      
max(sums_Archaea) #225892
min(sums_Archaea) # 68
sums_Archaea
#https://david-barnett.github.io/microViz/articles/web-only/tax-fixing.html #fixing the table
#https://rpubs.com/bioguo/680490 (Random forest classifications)
library(devtools);packageVersion("devtools")# ‘2.4.5’# Load the devtools package
#install_github("guokai8/microbial") # Install the package (Microbial package)
library(microbial);packageVersion("microbial")#‘0.0.22’####
ps_RandomForest_Archaea=Archaea
save(ps_RandomForest_Archaea,
file = "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps_RandomForest_Archaea.RData")


#Archaeal claasifier
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps_RandomForest_Archaea.RData")

prunescale = 0.05

# Prune out rare OTUs by mean relative abundance set by prunescale
tax.mean.ps_RandomForest_Archaea <- taxa_sums(ps_RandomForest_Archaea)/nsamples(ps_RandomForest_Archaea) # average number of reads for each otu
sites.prune.ps_RandomForest_Archaea <- prune_taxa(tax.mean.ps_RandomForest_Archaea > prunescale, ps_RandomForest_Archaea)

#Next, we are going to format a dataframe of predictors (OTUs) and responses (States)
# Make a dataframe of training data with OTUs as column and samples as rows
predictors.ps_RandomForest_Archaea<- t(otu_table(sites.prune.ps_RandomForest_Archaea));predictors.ps_RandomForest_Archaea
dim(predictors.ps_RandomForest_Archaea)
# 1197 1655

# Make one column for our outcome/response variable 
response.ps_RandomForest_Archaea <- as.factor(sample_data(sites.prune.ps_RandomForest_Archaea)$Genotype);response.ps_RandomForest_Archaea
# Combine them into 1 data frame
rf.data.ps_RandomForest_Archaea <- data.frame(response.ps_RandomForest_Archaea, predictors.ps_RandomForest_Archaea);rf.data.ps_RandomForest_Archaea
str(rf.data.ps_RandomForest_Archaea)
# Set a random seed
set.seed(5000)
# My computer has 8 processor cores - adjust parameter of the following command as needed/desired
registerDoParallel(3)

# %dopar% implemented with foreach implements each loop call as an independent function call
# and then parallelizes all the function calls across the number of processers set above with
# registerDoParallel(). First, we need to set up a dataframe to store each bootstrap RF result.
boot.imp.s <- data.frame(predictors=NULL, Try=NULL, MeanDecreaseGini=NULL)
boot.imp.s <- foreach (Try=1:100, .combine=rbind) %dopar% {
  
  ps_RandomForest_Archaea.classify <- randomForest(response.ps_RandomForest_Archaea ~., data = rf.data.ps_RandomForest_Archaea, ntree = 500, proximity=T, mtry=200)
  print(ps_RandomForest_Archaea.classify)
  
  # Make a data frame with predictor names and their importance
  imp.s <- importance(ps_RandomForest_Archaea.classify)
  imp.s <- data.frame(predictors = rownames(imp.s), imp.s)
  
  cbind(imp.s, data.frame(Try=Try)) #)
  
}

# Same (and independent) as above but this time we are storing results from the
# out-of-bag error rates, so we car report the correct-incorrect classification
# rate for each group, and what types of misclassifications were made.
boot.oob <- NULL
boot.oob <- foreach (Try=1:100, .combine=rbind) %dopar% {
  
  ps_RandomForest_Archaea.classify <- randomForest(response.ps_RandomForest_Archaea ~., data = rf.data.ps_RandomForest_Archaea, ntree = 500, proximity=T, mtry=200)
  print(ps_RandomForest_Archaea.classify)
  
  # Make a data frame with predictor names and their importance
  imp.s <- importance(ps_RandomForest_Archaea.classify)
  imp.s <- data.frame(predictors = rownames(imp.s), imp.s)
  
  c(boot.oob, ps_RandomForest_Archaea.classify$err.rate[500,1])
}

#Display and visualize results
# Look at the results
mean(boot.oob)#0.1007352
sd(boot.oob)#0.003860804
range(boot.oob)# 0.09106099 0.11027569
# Create a dataframe with summary statistics from our bootstrap runs
summary.boot.s <- summaryBy(MeanDecreaseGini ~ predictors, data= boot.imp.s, FUN = c(mean, sd, std.error));summary.boot.s
write.csv(summary.boot.s,"summary.boot.s.csv")
# Look at it
head(summary.boot.s)
str(summary.boot.s)
# Order the predictor levels by importance
imp.s.sort <- arrange(summary.boot.s, desc(MeanDecreaseGini.mean));imp.s.sort
imp.s.sort$predictors <- factor(imp.s.sort$predictors, levels = imp.s.sort$predictors)
# Select the top 50 predictors
imp.s.5 <- imp.s.sort[1:20, ];imp.s.5

# What are those OTUs?
# otunames.s = otunames of predictors
# r.s = A logical containing which rows from phyloseq object were recovered as top 40 predictors
otunames.s <- imp.s.5$predictors;otunames.s
r.s <- rownames(tax_table(ps_RandomForest_Archaea)) %in% otunames.s;r.s

# Use r.s to pull out the taxonomic information for those OTUS in otunames.s
t.table <- as.data.frame(tax_table(ps_RandomForest_Archaea)[r.s, ])
write.csv(t.table, "RF.ps_RandomForest_Archaea.t.table.csv", quote=F,row.names=T)


#VISUALIZATION HERE
# Reformat the taxonomic names and paste them to OTU names for axis labels
# (highest determined taxonomic rate, get rid of special characters, etc.)
axislabels <- with(t.table, paste(rownames(t.table), gsub("[sgfocp]__","",Genus,perl=T) %>% gsub("_"," ",.,perl=T) %>% gsub("[\\s]?(unclassified|ge)[\\s]?","",.,perl=T)))
names(axislabels) <- rownames(t.table)
i <- grep("(uncultured)|([01234567890-]$)+", axislabels, perl=T)
axislabels[i]<-paste(names(axislabels)[i], t.table$Family[i])

# Now we sum the abundance of each OTU by state so we can order the
# RF predictor OTUs by which state they were most abundant in, followed
# by their importance value. This makes the heatmap easier to interpret.
sums<-summaryBy(. ~ response.ps_RandomForest_Archaea, data = rf.data.ps_RandomForest_Archaea[, c('response.ps_RandomForest_Archaea', as.character(otunames.s))], FUN = sum, keep.names = T);sums
state.organize <- sapply(as.character(otunames.s), FUN = function(x) which.max(sums[,x]));state.organize
sums$response.ps_RandomForest_Archaea
rf.data.ps_RandomForest_Archaea$response.ps_RandomForest_Archaea
# hm is a dataframe of the abundance of the predictors, standardized and log transformed
hm <- rf.data.ps_RandomForest_Archaea[,as.character(otunames.s)] %>% decostand(method='max', MARGIN=2) %>% decostand(method='log');hm
hm$sample <- gsub('_',' ',sample_data(ps_RandomForest_Archaea)[rownames(hm),'Genotype']$Genotype)

#hm.melt <- melt(hm[c(order(state.organize), 6)]);hm.melt
hm.melt <-  as.data.table(hm[c(order(state.organize), 21)]);hm.melt
hm.melt <- melt(hm.melt, id.vars = "sample")
names(hm.melt) <- c("Tree","Taxon","StandardizedAbundance")

# Turn predictors into a factor (re-ordered the same as the hm.melt object so they match up)
imp.s.5$predictors <- factor(imp.s.5$predictors, levels(imp.s.5$predictors)[order(state.organize)])


# Creat the heatmap with the taxon names and OTUs
plot1_Archaea <- ggplot(hm.melt, aes(x=Tree, y=Taxon, fill=StandardizedAbundance)) + 
  geom_tile() + scale_fill_gradient(low="white", high="grey35") + 
  theme(axis.text.x = element_text(angle=90, size=14), legend.position = "none", axis.text.y = element_text(hjust=0))  + 
  scale_y_discrete(labels=axislabels)+theme(axis.title.x = element_text(face="bold", colour="black", size=14),
                                            axis.title.y = element_text(face="bold", colour="black", size=14),
                                            axis.text.x = element_text(face="bold", colour="black", size=14),
                                            axis.text.y  = element_text(face="bold", colour="black", size=14))+ labs(tag = "",size=14);plot1_Archaea

#Creat the bargraph
plot2_Archaea <- ggplot(imp.s.5, aes(x = predictors, y = MeanDecreaseGini.mean)) +
  geom_bar(stat = "identity", fill = "#7E6148B2") +
  ylab("Bootstrap Mean \u00b1 SD Gini Index") +
  theme_bw() +
  geom_errorbar( aes(x=predictors, ymin=MeanDecreaseGini.mean-MeanDecreaseGini.sd, ymax=MeanDecreaseGini.mean+MeanDecreaseGini.sd), width=0.75, alpha=0.9, size=0.5) +
  coord_flip() + theme(axis.title.x = element_text(face="bold", colour="black", size=14),
                       axis.title.y = element_text(face="bold", colour="black", size=14),
                       axis.text.x = element_text(face="bold", colour="black", size=14),
                       axis.text.y  = element_text(face="bold", colour="black", size=14))+ labs(tag = "",size=14);plot2_Archaea
g2_Archaea <- ggplotGrob(plot2_Archaea);g2_Archaea

library(ggpubr)
library(patchwork)

#Put them together

tiff("RF1_Archaea.tiff", units="in", width=15, height=6, res=300)
RF1_Archaea.tiff=plot1_Archaea+plot2_Archaea+plot_layout(nrow=1);RF1_Archaea.tiff
ggsave("RF1_Archaea.tiff", plot = RF1_Archaea.tiff)
dev.off()
```





Figures combined
#Rarefaction
```{r}
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
Rarefaction.tiff <- Rarefaction_Bacteria + Rareaction_Fungi + Rareaction_Archaea +
  plot_layout(nrow = 1, byrow = FALSE)

ggsave(
  "Rarefaction.tiff",
  plot = Rarefaction.tiff,
  device = "tiff",
  width = 9,
  height = 3.5,
  units = "in",
  dpi = 300
)
```


#Alpha_diversity
```{r}
tiff("Alpha_diversity.tiff", units="in", width=9, height=5, res=300)
Alpha_diversity.tiff=(plot_bacterial_rich+plot_bacterial_Shannon+plot_Fungi_rich+plot_Fungi_Shannon+plot_Archaea_rich+plot_Archaea_Shannon+plot_layout(nrow = 2, byrow = FALSE));Alpha_diversity.tiff
dev.off()
```


#Microbial_anbundance
```{r}
tiff("Abundance.tiff", units="in", width=9, height=3, res=300)
Abundance.tiff=(plot_bacterial_Abundance+plot_Fungil_Abundance+plot_Archaea_Abundance+plot_layout(nrow = 1, byrow = FALSE));Abundance.tiff
dev.off()
```


#Beta diversity
```{r}
tiff("Beta_diversity.tiff.tiff", units="in", width=9, height=5, res=300)
Beta_diversity.tiff.tiff=(pcaplot1_bacteria.tiff+pcoaplot1_bacteria.tiff+pcaplot1_fungi.tiff+pcoaplot1_fungi.tiff+pcaplot1_Archaea.tiff+pcoaplot1_Archaea.tiff+plot_layout(nrow = 2, byrow = FALSE));Beta_diversity.tiff.tiff
dev.off()
```

#core Microbiome 
```{r}
tiff("Core_Microbiome.tiff", units="in", width=9, height=5, res=300)
Core_Microbiome.tiff=(p_core_flex1_all_bacteria+p_core_flex1_all_Fungi+plot_layout(nrow = 1, byrow = FALSE));Core_Microbiome.tiff
dev.off()
```


#core Microbiome archaea
```{r}
tiff("Core_Microbiome_archaea.tiff", units="in", width=9, height=5, res=300)
Core_Microbiome_archaea.tiff=(p_core_flex1_all_Archaea+plot_layout(nrow = 1, byrow = FALSE));Core_Microbiome_archaea.tiff
dev.off()
```





#Correlation
```{r}
#Extract Bacterial Diversity
#To extract bacterial diversity, you can subset the phyloseq object by taxonomy level (e.g., Kingdom = #"Bacteria").
bacteria_ps <- ps3_Bacteria
bacteria_diversity <- estimate_richness(bacteria_ps, measures = c("Shannon", "Observed"));bacteria_diversity

#Extract Fungal Diversity
#Similarly, extract fungal diversity using a filter for Kingdom = "Fungi".

fungi_ps <- bacteria_ps <- ps3_Fungi
fungal_diversity <- estimate_richness(fungi_ps, measures = c("Shannon", "Observed"));fungal_diversity

#3. Merge the Diversity Data
#Next, combine the bacterial and fungal diversity data. Assuming both datasets have the same sample IDs, you can merge them:

diversity_data <- cbind(bacteria_diversity, fungal_diversity);diversity_data
colnames(diversity_data)[1] <- "Bacteria_richness"
colnames(diversity_data)[2] <- "Bacteria_Shannon"
colnames(diversity_data)[3] <- "Fungal_richness"
colnames(diversity_data)[4] <- "Fungal_Shannon"
diversity_data

write.csv(diversity_data, "diversity_data.csv")
```

#correlate richness

```{r}
#https://bookdown.org/content/b298e479-b1ab-49fa-b83d-a57c2b034d49/correlation.html#correlogram
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
dat=read.csv("diversity_data.csv", header = T);dat


# basic scatterplot
corr1=ggplot(dat, aes(x=Bacteria_richness, y=Fungal_richness, alpha=Genotype)) + 
    geom_point(size=6, color="#69b3a2") +
    theme_ipsum();corr1

tiff("corr1.tiff", units="in", width=9, height=5, res=300)
corr1.tiff=(corr1+plot_layout(nrow = 1, byrow = FALSE));corr1.tiff
dev.off()
```


```{r}
library(GGally)
# Check correlation between variables
#cor(dat)
ggcorr(dat) 
 
# Nice visualization of correlations
 
# Check correlations (as scatterplots), distribution and print corrleation coefficient 
corr2=ggpairs(dat, columns = 2:5, ggplot2::aes(colour=Genotype));corr2  # Reduce overall text size

 ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/corr2.tiff",     
  plot = corr2,
   scale = 1,
 width = 25,
  height = 20,
  units = c("cm"),
 dpi = 600)

```





```{r}
#https://rpubs.com/Alema/1000474
setwd("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome")
dat=read.csv("diversity_data.csv", header = T);dat
library(GGally)
corr3=ggpairs(dat, columns = 2:5, title="correlogram with ggpairs()");corr3


 ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/corr3.tiff",     
  plot = corr3,
   scale = 1,
 width = 25,
  height = 20,
  units = c("cm"),
 dpi = 600)
```


#Preleminary test to check the test assumptions
#data are normally distributed
#Is the covariation linear? Yes, form the plot, the relationship is linear. In the situation where the scatter plots show curved patterns, we are dealing with nonlinear association between the two variables.
#Are the data from each of the 2 variables (x, y) follow a normal distribution?

#Use Shapiro-Wilk normality test -> R function: shapiro.test() and look at the normality plot -> R function: ggpubr::ggqqplot()

#Shapiro-Wilk test can be performed as follow:

#Null hypothesis: the data are normally distributed Alternative hypothesis: the data are not normally distributed
```{r}
#install.packages("ggpubr")
library(ggpubr)

shapiro.test(dat$Bacteria_richness)
#data:  dat$Bacteria_richness
#W = 0.76335, p-value = 1.508e-05

shapiro.test(dat$Bacteria_Shannon)
#data:  dat$Bacteria_Shannon
#W = 0.86125, p-value = 0.001077

shapiro.test(dat$Fungal_richness)
#W = 0.91085, p-value = 0.01563

shapiro.test(dat$Fungal_Shannon)
#W = 0.91628, p-value = 0.02149
```


#data are not normally distributed
#If the data are not normally distributed, it’s recommended to use the non-parametric correlation, including Spearman and Kendall rank-based correlation tests.
```{r}
#Spearman rank correlation coefficient
cor.test(dat$Bacteria_richness, dat$Fungal_richness,  method = "spearman")

#Warning: Cannot compute exact p-value with ties
#	Spearman's rank correlation rho

#data:  dat$Bacteria_richness and dat$Fungal_richness
#S = 1550.9, p-value = 8.583e-05
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#      rho 
#0.6549767 

cor.test(dat$Bacteria_Shannon, dat$Fungal_Shannon,  method = "spearman")

#	Spearman's rank correlation rho

#data:  dat$Bacteria_Shannon and dat$Fungal_Shannon
#S = 944, p-value = 1.45e-06
#alternative hypothesis: true rho is not equal to 0
#sample estimates:
#      rho 
#0.7899889  
library(ggpubr)
sp <- ggscatter(dat, x = "Bacteria_richness", y = "Fungal_richness",
   color = "Genotype", palette = "jco",
   add = "reg.line", conf.int = TRUE)
sp + stat_cor(aes(color = Genotype), label.x = 3)

```


#Functional studies using the microco packages

#using the microeco package, we use the file convention there to compute for the assignment of microbiome functions of the seed bacterial community
##https://chiliubio.github.io/microeco_tutorial/explainable-class.html
#https://chiliubio.github.io/microeco_tutorial/basic-class.html#microtable-class
# Dataset objects

```{r}
#Load packages 
#Load the functions
library(microeco)
library(magrittr)
library(ggplot2)
library(aplot)
library(phyloseq)
source("phylo-meco.txt")
library(ggh4x)
library(ggradar)
library(ggnested)
library(circlize)  # to create continuous color gradients
library(dplyr)  # wrangling tidy tables
library(grid)  # to define witdth / height of annotations using grid::unit()
library(ragg)  # to generate high quality graphics
library(RColorBrewer)  # to access predefined color palettes
library(tibble)  # to convert data.frames to tibbles
library(tidyHeatmap) # the main act: plotting & annotating heatmaps
library(readr)
library(edgeR)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ape)
library(plyr)
library(gplots)
library(lme4)
library(tidyr)
library(vegan)
library(scales)
library(gridExtra)
library(grid)
library(reshape2)
library(phyloseq)
library(stringr)
library(compositions)
library(zCompositions)

##Side comment: Do I need all these packages? No, probably not. But it is more convenient generaly to copy and paste most of these packages into every document.


m.g.colors <- c( "#999999", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861" );m.g.colors

theme_set(theme_classic(base_size = 14))

#Bacteria
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Bacteria.RData")

#Fungi
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")


```
#Convert to phyloseq object
#ps_Bacteria <- merge_phyloseq(
#  asv_tab_bac,
# tax_tab_bac, 
# metadata_bac);ps_Bacteria#[ 16460 taxa & 1197 samples ]
#summarize_phyloseq(ps_Bacteria)


#Bacterial community
```{r}
#object=ps3_Bacteria
taxdf = data.frame(tax_table(ps3_Bacteria))
taxdf[1:4,1:4]
colnames(taxdf) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")

#FAPROTAX wants one string with all the column names.
taxdf$taxstring = paste(taxdf$Kingdom, taxdf$Phylum, taxdf$Class, taxdf$Order, taxdf$Family, taxdf$Genus, sep="; ")
taxdf[1:4,]

#We’re now appending the taxonomy string as a column to the otu table. Save the otu table as a .tsv file and tell R to use a tab character as the separater. Quotes around the values will break FAPROTAX.
otutable=data.frame(otu_table(ps3_Bacteria));otutable
otutable$taxonomy = taxdf[as.vector(row.names(otutable)),"taxstring"]
otutable[1:4,]

write.table(otutable, "faprotax.taxon.table.tsv", sep = "\t", quote=FALSE)
```

#https://mgaley-004.github.io/MiSeq-Analysis/Tutorials/faprotax_demo.html
#Run FAPROTAX
#FAPROTAX is a command line utility that relys on having python 3 installed on your system. I personally have had dependency problems running FAPROTAX with python 3.9 and recommend using python 3.7.

#In terminal or a shell of your choice, navigate to the directory where you have downloaded and unpacked FAPROTAX. Handy reminders on UNIX vs MsDOS file system navigation.

#If you are running MacOS or a Linux based OS:

```{r}
#./collapse_table.py -i faprotax.taxon.table.tsv -o out.functable.tsv -g FAPROTAX.txt -c '#' -d taxonomy --omit_columns 0 --column_names_are_in last_comment_line -r out.report.txt -v -s out_subtables/ -f

#If you are running Windows:
#python collapse_table.py -i faprotax.taxon.table.tsv -o out.functable.tsv -g FAPROTAX.txt -c '#' -d taxonomy --omit_columns 0 --column_names_are_in last_comment_line -r out.report.txt -v -s out_subtables/ -f
```



#You may recieve many messages warning you that ‘is’ != ==, but don’t worry. It’s working.

#An explanation of all of the flags we used:

#./collapse_table.py or python collapse_table.py runs the main faprotax program
#-i faprotax.taxon.table.tsv this is the argument for the name of your input file.
#-o out.functable.tsv this is the argument for the name of the output file you want.
#-g FAPROTAX.txt this is the name of the faprotax database. don’t change this.
#-c ‘#’ this tells faprotax that ‘#’ is the comment character. don’t change.
#-d taxonomy this tells faprotax to look for taxonomy information in the column named ‘taxonomy’. you would only change this if you changed the name of the column in the R code.
#–omit_columns 0 this tells faprotax to ignore the first column, which is just OTU names.
#–column_names_are_in_last_comment_line this tells faprotax where the sample names are. don’t change.
#-r out.report.txt this gives faprotax the name for the report file to create.
#-v this makes visible output on your screen while faprotax is working. I find it reassuring.
#-s out_subtables/ this tells faprotax to separate out and save otu_tables for each functional groups and store them in a folder called out_subtables. You don’t need to make that folder first, faprotax will do it for you.
#-f this is optional. It tells faprotax that it’s okay to overwrite filenames. I put this in there when I was troubleshooting this script and didn’t want to keep making new file names.
#You can read additional documentation on how to run FAPROTAX on their website.


#FAPROTAX results
#Broad patterns
#First we’ll look at the major functions FAPROTAX identified in the samples regardless of taxonomy.

#You must remove the comment symbol from the first line of the results tsv for the following to work.


```{r}
otu.function.table <- read.csv("FAPROTAX_1.2.10/out.functable.csv",
              header=TRUE,
              row.names=1);otu.function.table
#otu.function.table <- read.table("FAPROTAX_1.2.10/out.functable.tsv", sep="\t", header=TRUE, row.names=1);otu.function.table

#Discard functions identified in fewer than 10 reads
otu.function.table <- otu.function.table[which(rowSums(otu.function.table) > 10),]
#Discard samples where no functions were identified
otu.function.table <- otu.function.table[,which(colSums(otu.function.table) > 1)]
otu.function.table[1:4,1:4]
```


```{r}
#Check how many functions were identified
dim(otu.function.table)

#This normalizes the read counts to the total function identifications for each samples. An imperfect solution since FAPROTAX most likely did not assign a function to all of your OTUS. To check FAPROTAX's performance, read the report file.

rel.otu.fn.table <- apply(otu.function.table,MARGIN = 2, FUN = function(x) x/sum(x))

rel.otu.fn.df <- data.frame(rel.otu.fn.table);rel.otu.fn.df
```



#Load metadata
```{r}
sampledata = read.csv("FAPROTAX_1.2.10/Bacteria_metadata.csv", row.names=1, header=TRUE);sampledata
#Convert the results dataframe into a longform dataframe (melt it).
```



#Convert the results dataframe into a longform dataframe (melt it).
```{r}
#Make row names a column
rel.otu.fn.df$BioFunction = row.names(rel.otu.fn.df)
rel.otu.fn.df$BioFunction
#melt
library(data.table)

# convert to data.table
otu.melt <- melt(setDT(rel.otu.fn.df), id.vars = "BioFunction");otu.melt
#otu.melt <- melt(rel.otu.fn.df, id.vars = "BioFunction")

#make this a sparse dataframe
otu.s.melt <- otu.melt[which(otu.melt$value > 0),];otu.s.melt
```

# Pull relevant metadata and add to the melted dataframe.
```{r}

colnames(otu.s.melt) <- c("BioFunction", "variable", "value")
otu.s.melt$Genotype <- sampledata[as.vector(otu.s.melt$variable),"Genotype"]
otu.s.melt$Sample_type <- sampledata[as.vector(otu.s.melt$variable),"Sample_type"]
otu.s.melt$target <- sampledata[as.vector(otu.s.melt$variable),"target"]
otu.s.melt$Barcode_system <- sampledata[as.vector(otu.s.melt$variable),"Barcode_system"]
#otu.s.melt$SA_code <- sampledata[as.vector(otu.s.melt$variable),"SA_code"]
#otu.s.melt$wr.dens.cat <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.cat"]
#otu.s.melt$wr.dens.scale <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.scale"]

otu.s.melt[1:10,]
```

#The Problem
```{r}
theme_set(theme_classic(base_size=12))

Relative_bacterial_func=ggplot(otu.s.melt[which(otu.s.melt$value >= 0.005),], aes(x=Genotype, y=value, fill=BioFunction)) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values=c(mycolors, "pink"));Relative_bacterial_func

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Relative_bacterial_func.tiff",     
  plot = Relative_bacterial_func,
   scale = 1,
 width = 11,
  height = 9,
  units = c("cm"),
 dpi = 600)

```

#All this kind of chart can tell us about is how frequently OTUs were assigned a function out of all possible assignments. It’s important to remember that FAPROTAX does not create a compositional dataset! OTUs are assigned to multiple categories. One obvious example is that ‘aerobic chemoheterotrophs’ are necessarily ‘chemoheterotrophs’, so the fact that the blue and gray bars take up most of the space in the chart together is really misleading. Many of the other assignments also overlap. For example, many of the identified nitrate reducing bacteria also happen to be iron respiring bacteria and chemoheterotrophs.

#That is why stacked bar charts and any kind of chart that implies a compositional set are inappropriate for this kind of data.


#Solution: Relative Scaling
#This will not solve the bar chart problem, but will help to make the data itself more interpretable. Instead of normalizing the count values to the sum of the faprotax table, normalize them to the total read depth.



```{r}
#rel.otu.fn.table <- sapply(colnames(otu.function.table), function(x) otu.function.table[,x]/raw.depth[x]);rel.otu.fn.table

raw.depth <- colSums(otu.function.table);raw.depth
rel.otu.fn.table <- sweep(otu.function.table, 2, raw.depth, "/");rel.otu.fn.table

rel.otu.fn.df <- data.frame(rel.otu.fn.table);rel.otu.fn.df
row.names(rel.otu.fn.df) <- row.names(otu.function.table)

rel.otu.fn.df[1:10,1:4]
```

#melt the data again
```{r}
rel.otu.fn.df$BioFunction = row.names(rel.otu.fn.df)
#melt
library(data.table)

# convert to data.table
otu.melt <- melt(setDT(rel.otu.fn.df), id.vars = "BioFunction");otu.melt
#otu.melt <- melt(rel.otu.fn.df, id.vars = "BioFunction")

#make this a sparse dataframe
otu.s.melt <- otu.melt[which(otu.melt$value > 0),];otu.s.melt
```



```{r}
colnames(otu.s.melt) <- c("BioFunction", "variable", "value")
otu.s.melt$Genotype <- sampledata[as.vector(otu.s.melt$variable),"Genotype"]
otu.s.melt$Sample_type <- sampledata[as.vector(otu.s.melt$variable),"Sample_type"]
otu.s.melt$target <- sampledata[as.vector(otu.s.melt$variable),"target"]
otu.s.melt$Barcode_system <- sampledata[as.vector(otu.s.melt$variable),"Barcode_system"]
#otu.s.melt$SA_code <- sampledata[as.vector(otu.s.melt$variable),"SA_code"]
#otu.s.melt$wr.dens.cat <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.cat"]
#otu.s.melt$wr.dens.scale <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.scale"]

otu.s.melt[1:10,]

```



#Visualization
```{r}
otu.s.melt$BioFunction <- as.factor(otu.s.melt$BioFunction)

#Optionally, abbreviate the long names of the FAPROTAX functions to make your charts more readable. 

otu.s.melt$ShortFunction <- otu.s.melt$BioFunction
#levels(otu.s.melt$ShortFunction) <- c("ONH3O", "OCHT", "ONO3O", "ANMHD", "APSY", "AOPHAT", "AOPHATS", "ARCD", "ARHD", "CELL", "CHT", "CHIT", "CLRE", "CYAN", "DH2O", "DFEO", "DSCO", "DSO", "DENI", "FERM", "FURE", "HUAS", "HUGU", "HUPA", "HCDE", "INPA", "FERE", "MAGU", "CHO", "CHT", "CHLT", "NO3DE", "NO3RE", "NO3RS", "NO2F", "NO2DEN", "NO2RS", "N2FX", "N2RS", "NODEN", "OPHAT", "PHAT", "PHHT", "PHT", "PRED", "REDACT", "SCRES", "SO4RES", "SO3RES", "SRES", "TSRES", "UREO")
```



#Heat maps! They don’t imply a composition and also let you off the hook for picking colors.
```{r}
# Choose a smaller font size
theme_set(theme_classic(base_size=8))

sorted.melt <- otu.s.melt
faprotax_bysample=ggplot(sorted.melt[which(sorted.melt$value >= 0.02),], aes(x=variable, y=tolower(ShortFunction), fill=value)) + geom_tile() + scale_fill_viridis_c(limits=c(0,0.3)) + theme(axis.text.x = element_blank()) + ylab("FAPROTAX Function") + facet_wrap(~Genotype, scales = "free_x");faprotax_bysample
```

#Save your work!
```{r}
ggsave("faprotax_bysample.tiff",width=7, height=5, units="in", dpi=300) 
```








#Archaea community
#load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")

```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Archaea.RData")
#object=ps3_Archaea
taxdf = data.frame(tax_table(ps3_Archaea))
taxdf[1:4,1:4]
colnames(taxdf) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")

#FAPROTAX wants one string with all the column names.
taxdf$taxstring = paste(taxdf$Kingdom, taxdf$Phylum, taxdf$Class, taxdf$Order, taxdf$Family, taxdf$Genus, sep="; ")
taxdf[1:4,]

#We’re now appending the taxonomy string as a column to the otu table. Save the otu table as a .tsv file and tell R to use a tab character as the separater. Quotes around the values will break FAPROTAX.
otutable=data.frame(otu_table(ps3_Archaea));otutable
otutable$taxonomy = taxdf[as.vector(row.names(otutable)),"taxstring"]
otutable[1:4,]

write.table(otutable, "faprotax.taxon.table.archaea.tsv", sep = "\t", quote=FALSE)
```


#Run FAPROTAX
#FAPROTAX is a command line utility that relys on having python 3 installed on your system. I personally have had dependency problems running FAPROTAX with python 3.9 and recommend using python 3.7.

#In terminal or a shell of your choice, navigate to the directory where you have downloaded and unpacked FAPROTAX. Handy reminders on UNIX vs MsDOS file system navigation.

#If you are running MacOS or a Linux based OS:

```{r}
#./collapse_table.py -i faprotax.taxon.table.archaea.tsv -o out.functable.tsv -g FAPROTAX.txt -c '#' -d taxonomy --omit_columns 0 --column_names_are_in last_comment_line -r out.report.txt -v -s out_subtables/ -f

#If you are running Windows:
#python collapse_table.py -i faprotax.taxon.table.archaea.tsv -o out.functable.tsv -g FAPROTAX.txt -c '#' -d taxonomy --omit_columns 0 --column_names_are_in last_comment_line -r out.report.txt -v -s out_subtables/ -f
```



#You may recieve many messages warning you that ‘is’ != ==, but don’t worry. It’s working.

#An explanation of all of the flags we used:

#./collapse_table.py or python collapse_table.py runs the main faprotax program
#-i faprotax.taxon.table.archaea.tsv this is the argument for the name of your input file.
#-o out.functable.tsv this is the argument for the name of the output file you want.
#-g FAPROTAX.txt this is the name of the faprotax database. don’t change this.
#-c ‘#’ this tells faprotax that ‘#’ is the comment character. don’t change.
#-d taxonomy this tells faprotax to look for taxonomy information in the column named ‘taxonomy’. you would only change this if you changed the name of the column in the R code.
#–omit_columns 0 this tells faprotax to ignore the first column, which is just OTU names.
#–column_names_are_in_last_comment_line this tells faprotax where the sample names are. don’t change.
#-r out.report.txt this gives faprotax the name for the report file to create.
#-v this makes visible output on your screen while faprotax is working. I find it reassuring.
#-s out_subtables/ this tells faprotax to separate out and save otu_tables for each functional groups and store them in a folder called out_subtables. You don’t need to make that folder first, faprotax will do it for you.
#-f this is optional. It tells faprotax that it’s okay to overwrite filenames. I put this in there when I was troubleshooting this script and didn’t want to keep making new file names.
#You can read additional documentation on how to run FAPROTAX on their website.


#FAPROTAX results
#Broad patterns
#First we’ll look at the major functions FAPROTAX identified in the samples regardless of taxonomy.

#You must remove the comment symbol from the first line of the results tsv for the following to work.


```{r}
otu.function.table <- read.csv("FAPROTAX_1.2.10/out.functable.arc.csv",
              header=TRUE,
              row.names=1);otu.function.table
#otu.function.table <- read.table("FAPROTAX_1.2.10/out.functable.tsv", sep="\t", header=TRUE, row.names=1);otu.function.table

#Discard functions identified in fewer than 10 reads
otu.function.table <- otu.function.table[which(rowSums(otu.function.table) > 10),]
#Discard samples where no functions were identified
otu.function.table <- otu.function.table[,which(colSums(otu.function.table) > 1)]
otu.function.table[1:4,1:4]
```


```{r}
#Check how many functions were identified
dim(otu.function.table)

#This normalizes the read counts to the total function identifications for each samples. An imperfect solution since FAPROTAX most likely did not assign a function to all of your OTUS. To check FAPROTAX's performance, read the report file.

rel.otu.fn.table <- apply(otu.function.table,MARGIN = 2, FUN = function(x) x/sum(x))

rel.otu.fn.df <- data.frame(rel.otu.fn.table);rel.otu.fn.df
```



#Load metadata
```{r}
sampledata = read.csv("FAPROTAX_1.2.10/Archaea_Metadata.csv", row.names=1, header=TRUE);sampledata
#Convert the results dataframe into a longform dataframe (melt it).
```



#Convert the results dataframe into a longform dataframe (melt it).
```{r}
#Make row names a column
rel.otu.fn.df$BioFunction = row.names(rel.otu.fn.df)
rel.otu.fn.df$BioFunction
#melt
library(data.table)

# convert to data.table
otu.melt <- melt(setDT(rel.otu.fn.df), id.vars = "BioFunction");otu.melt
#otu.melt <- melt(rel.otu.fn.df, id.vars = "BioFunction")

#make this a sparse dataframe
otu.s.melt <- otu.melt[which(otu.melt$value > 0),];otu.s.melt
```

# Pull relevant metadata and add to the melted dataframe.
```{r}

colnames(otu.s.melt) <- c("BioFunction", "variable", "value")
otu.s.melt$Genotype <- sampledata[as.vector(otu.s.melt$variable),"Genotype"]
otu.s.melt$Sample_type <- sampledata[as.vector(otu.s.melt$variable),"Sample_type"]
otu.s.melt$target <- sampledata[as.vector(otu.s.melt$variable),"target"]
otu.s.melt$Barcode_system <- sampledata[as.vector(otu.s.melt$variable),"Barcode_system"]
#otu.s.melt$SA_code <- sampledata[as.vector(otu.s.melt$variable),"SA_code"]
#otu.s.melt$wr.dens.cat <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.cat"]
#otu.s.melt$wr.dens.scale <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.scale"]

otu.s.melt[1:10,]
```

#The Problem
```{r}
theme_set(theme_classic(base_size=12))

Relative_archaeal_func=ggplot(otu.s.melt[which(otu.s.melt$value >= 0.005),], aes(x=Genotype, y=value, fill=BioFunction)) + geom_bar(stat="identity", position="fill") + scale_fill_manual(values=c(mycolors, "pink"));Relative_archaeal_func

ggsave(
   "/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/Relative_archaeal_func.tiff",     
  plot = Relative_archaeal_func,
   scale = 1,
 width = 13,
  height = 9,
  units = c("cm"),
 dpi = 600)

```

#All this kind of chart can tell us about is how frequently OTUs were assigned a function out of all possible assignments. It’s important to remember that FAPROTAX does not create a compositional dataset! OTUs are assigned to multiple categories. One obvious example is that ‘aerobic chemoheterotrophs’ are necessarily ‘chemoheterotrophs’, so the fact that the blue and gray bars take up most of the space in the chart together is really misleading. Many of the other assignments also overlap. For example, many of the identified nitrate reducing bacteria also happen to be iron respiring bacteria and chemoheterotrophs.

#That is why stacked bar charts and any kind of chart that implies a compositional set are inappropriate for this kind of data.


#Solution: Relative Scaling
#This will not solve the bar chart problem, but will help to make the data itself more interpretable. Instead of normalizing the count values to the sum of the faprotax table, normalize them to the total read depth.



```{r}
#rel.otu.fn.table <- sapply(colnames(otu.function.table), function(x) otu.function.table[,x]/raw.depth[x]);rel.otu.fn.table

raw.depth <- colSums(otu.function.table);raw.depth
rel.otu.fn.table <- sweep(otu.function.table, 2, raw.depth, "/");rel.otu.fn.table

rel.otu.fn.df <- data.frame(rel.otu.fn.table);rel.otu.fn.df
row.names(rel.otu.fn.df) <- row.names(otu.function.table)

rel.otu.fn.df[1:10,1:4]
```

#melt the data again
```{r}
rel.otu.fn.df$BioFunction = row.names(rel.otu.fn.df)
#melt
library(data.table)

# convert to data.table
otu.melt <- melt(setDT(rel.otu.fn.df), id.vars = "BioFunction");otu.melt
#otu.melt <- melt(rel.otu.fn.df, id.vars = "BioFunction")

#make this a sparse dataframe
otu.s.melt <- otu.melt[which(otu.melt$value > 0),];otu.s.melt
```



```{r}
colnames(otu.s.melt) <- c("BioFunction", "variable", "value")
otu.s.melt$Genotype <- sampledata[as.vector(otu.s.melt$variable),"Genotype"]
otu.s.melt$Sample_type <- sampledata[as.vector(otu.s.melt$variable),"Sample_type"]
otu.s.melt$target <- sampledata[as.vector(otu.s.melt$variable),"target"]
otu.s.melt$Barcode_system <- sampledata[as.vector(otu.s.melt$variable),"Barcode_system"]
#otu.s.melt$SA_code <- sampledata[as.vector(otu.s.melt$variable),"SA_code"]
#otu.s.melt$wr.dens.cat <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.cat"]
#otu.s.melt$wr.dens.scale <- sampledata[as.vector(otu.s.melt$variable),"wr.dens.scale"]

otu.s.melt[1:10,]

```



#Visualization
```{r}
otu.s.melt$BioFunction <- as.factor(otu.s.melt$BioFunction)

#Optionally, abbreviate the long names of the FAPROTAX functions to make your charts more readable. 

otu.s.melt$ShortFunction <- otu.s.melt$BioFunction
#levels(otu.s.melt$ShortFunction) <- c("ONH3O", "OCHT", "ONO3O", "ANMHD", "APSY", "AOPHAT", "AOPHATS", "ARCD", "ARHD", "CELL", "CHT", "CHIT", "CLRE", "CYAN", "DH2O", "DFEO", "DSCO", "DSO", "DENI", "FERM", "FURE", "HUAS", "HUGU", "HUPA", "HCDE", "INPA", "FERE", "MAGU", "CHO", "CHT", "CHLT", "NO3DE", "NO3RE", "NO3RS", "NO2F", "NO2DEN", "NO2RS", "N2FX", "N2RS", "NODEN", "OPHAT", "PHAT", "PHHT", "PHT", "PRED", "REDACT", "SCRES", "SO4RES", "SO3RES", "SRES", "TSRES", "UREO")
```



#Heat maps! They don’t imply a composition and also let you off the hook for picking colors.
```{r}
# Choose a smaller font size
theme_set(theme_classic(base_size=8))

sorted.melt <- otu.s.melt
faprotax_bysample.archaeal=ggplot(sorted.melt[which(sorted.melt$value >= 0.02),], aes(x=variable, y=tolower(ShortFunction), fill=value)) + geom_tile() + scale_fill_viridis_c(limits=c(0,0.3)) + theme(axis.text.x = element_blank()) + ylab("FAPROTAX Function") + facet_wrap(~Genotype, scales = "free_x");faprotax_bysample.archaeal
```

#Save your work!
```{r}
ggsave("faprotax_bysample.archaeal.tiff",width=7, height=5, units="in", dpi=300) 
```





#Fungal community
#Assign ecological guilds / functions
#Works with taxonomy (usually genus or species level).
#Matches taxa against a curated database of fungal lifestyles (e.g., Saprotroph, Pathotroph, Symbiotroph).
```{r}
load("/Users/expeditoolimi/Documents/Southampton/Tosin/African-seed-microbiome/ps3_Fungi.RData")
devtools::install_github("brendanf/FUNGuildR")

# Convert phyloseq to dataframe
taxonomy <- as.data.frame(tax_table(ps3_Fungi));taxonomy


# Match by genus or species
annotated <- taxonomy %>%
  left_join(traits, by = c("Species" = "Taxon"))

```

# Assigning Functional Guilds o terminal

```{r}
#Use -h or --help for a detail FUNGuild manual.

##Preparing your workspace:

#git clone https://github.com/UMNFuN/FUNGuild
#cd FUNGuild/
#Run the script using default parameters:

#python Guilds_v1.0.py -otu /Users/username/Documents/project/otu_table.txt -db fungi
```

#Then reaf the assigned functions

```{r}
# pseudocode - depends on package you install
df <- read.delim("FUNGuild/Fungi_Table.guilds.txt",, header = TRUE, sep = "\t", stringsAsFactors = FALSE);df
library(dplyr)
library(tidyr)
library(readr)

# Step 1: Filter by confidence
df_filtered <- df %>%
  filter(Confidence_Ranking %in% c("Probable", "Highly Probable"))

# Step 2: Identify sample columns
sample_cols <- grep("^f_[A-E][1-6]$", colnames(df_filtered), value = TRUE)

# Step 3: Pivot longer
df_long <- df_filtered %>%
  select(all_of(c("Guild", sample_cols))) %>%
  pivot_longer(
    cols = all_of(sample_cols),
    names_to = "Sample",
    values_to = "Abundance"
  )

# Step 4: Summarize by Sample and Guild
guild_summary <- df_long %>%
  group_by(Sample, Guild) %>%
  summarise(Total_Abundance = sum(Abundance, na.rm = TRUE), .groups = "drop")

# Step 5: Inspect result
head(guild_summary)


```



# Gather sample columns into long format
```{r}
library(ggplot2)
ggplot(guild_summary, aes(x = Sample, y = Total_Abundance, fill = Guild)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Fungal Guild Composition by Sample", x = "Sample", y = "Total Abundance")

```


